### 2018-01-02 paper/src/paper.tex
paper fix

\documentclass[10pt,journal,compsoc]{IEEEtran}

\usepackage{listings}
\usepackage{dirtree}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bbm}
\usepackage[scientific-notation=true]{siunitx}

\usepackage{amssymb}
\usepackage{tabulary}

\usepackage{etoolbox}

\usepackage{algorithm}
\usepackage{algorithmic}

\makeatletter
\newcommand\fs@norules{\def\@fs@cfont{\bfseries}\let\@fs@capt\floatc@ruled
  \def\@fs@pre{}%
  \def\@fs@post{}%
  \def\@fs@mid{\kern3pt}%
  \let\@fs@iftopcapt\iftrue}
\makeatother
\floatstyle{norules}
\restylefloat{algorithm}

\let\tinymatrix\smallmatrix
\let\endtinymatrix\endsmallmatrix
\patchcmd{\tinymatrix}{\scriptstyle}{\scriptscriptstyle}{}{}
\patchcmd{\tinymatrix}{\scriptstyle}{\scriptscriptstyle}{}{}
\patchcmd{\tinymatrix}{\vcenter}{\vtop}{}{}
\patchcmd{\tinymatrix}{\bgroup}{\bgroup\scriptsize}{}{}

\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

\usepackage[draft,index]{fixme}
\fxsetup{theme=color,mode=multiuser,layout=inline,draft}

% correct bad hyphenation here
%\hyphenation{op-tical net-works semi-conduc-tor}

%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10011007.10010940.10010971.10010980.10010986</concept_id>
%<concept_desc>Software and its engineering~Massively parallel systems</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%<concept>
%<concept_id>10010147.10010169</concept_id>
%<concept_desc>Computing methodologies~Parallel computing methodologies</concept_desc>
%<concept_significance>300</concept_significance>
%</concept>
%<concept>
%<concept_id>10010147.10010257.10010282.10010284</concept_id>
%<concept_desc>Computing methodologies~Online learning settings</concept_desc>
%<concept_significance>300</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc[500]{Software and its engineering~Massively parallel systems}
%\ccsdesc[300]{Computing methodologies~Parallel computing methodologies}
%\ccsdesc[300]{Computing methodologies~Online learning settings}

%\keywords{Multi-Armed Bandits, EASY-Backfilling, Parallel Job Scheduling, High
%Performance Computing, Simulation.}

%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Online (Bandit) Policy Selection for EASY-Backfilling}

\author{
\IEEEauthorblockN{Eric Gaussier}\\
\IEEEauthorblockA{Univ. Grenoble Alpes, CNRS, Grenoble INP, LIG\\
  eric.gaussier@imag.fr } \\ 
\IEEEauthorblockN{J\'er\^ome Lelong}\\
\IEEEauthorblockA{Univ. Grenoble Alpes, Inria, CNRS, Grenoble INP, LJK\\
jerome.lelong@imag.fr}  \\ 
\IEEEauthorblockN{Valentin Reis, Denis Trystram}\\
\IEEEauthorblockA{Univ. Grenoble Alpes, Inria, CNRS, Grenoble INP, LIG\\
denis.trystram@imag.fr}   
}
\begin{document}

\begin{abstract}

  The EASY-FCFS heuristic is the basic building block of job scheduling
  policies in most parallel High Performance Computing platforms. Despite its
  simplicity, and the guarantee of no job starvation, it could still be
  improved on a per-system basis. Such tuning is difficult because of
  non-linearities in the scheduling process. The study conducted in this paper
  considers an online approach to the automatic tuning of the EASY heuristic
  for HPC platforms.  More precisely, we consider the problem of selecting a
  reordering policy for the job queue under several feedback modes. We show via
  a comprehensive experimental validation on actual logs that noisy feedback
  (using a simulator, which is necessarily imprecise) recovers existing
  \textit{in-hindsight} results that allow to divide the average waiting time
  by almost 2. Moreover, we show that bandit feedback can be used by an
  Epsilon-greedy multi-armed bandit algorithm to cut the average waiting time
  by 40\% without using a simulator.

\end{abstract}

\maketitle

\section{Introduction}
\label{sec:intro}

\subsection{Context}

Providing the computational infrastuctures needed to solve actual complex
problems arising in the various fields of modern society (including climate
change, health, green energy or security) is a strategic challenge. The main
pillar to adress this problem is to build extreme-scale parallel and distributed
platforms. The never-ending race for more computing power and storage capacity
leads to sophisticated specific Exascale platforms, and the objective of the
community is to design efficient sustained Petascale platforms. This
large-scale evolution and the increasing complexity in both architecture and
applications create many scientific and technical problems.  Accordingly,
system management software still has a long way to go. The existing job and
resource management software allow to run tens of thousands of jobs on hundreds
of thousands of cores. Such software are based on robust policies with
positive properties such as relative simplicity and prevention of job
starvation. However, there is still room for improvement. Here, we study how to
tune the ordering of the submission queues in the frame of the classical
so-called EASY\footnote{Extensible Argonne Scheduling sYstem
(EASY)}-Backfilling family of heuristics using two new online approaches.  One
approach is based on a simulator and the second uses a multi-armed bandit
algorithm.

Most common work in the literature consider \textit{a posteriori} optimization
of a scheduling algorithm on a (actual) data set.  The increasing amount and
diversity of data generated by large scale platforms motivate the community to
study more sophisticated policies that could leverage these data.  During the
last few years, there was an explosion in the number of works at the interface
of the HPC and Big Data fields, dealing with learning algorithms. Recent
sophisticated approaches developed \textit{ad hoc} supervised machine learning
algorithms for reducing uncertainty in problem parameters (e.g.,\ job execution
times\cite{learningruntimes}, memory requirements, etc.). However, only few studies
evaluate the impact of such techniques on the performances of the resource
manager.

Unlike most existing studies, which consider the learning of specific
parameters of a trace, we provide algorithms that use feedback from the system
in order to alter the scheduling behavior online. Our focus is to learn how the
platform behaves within the usual framework of EASY-Backfilling described in
Section~\ref{sec:rw} . More precisely, we provide strategies for choosing the
most appropriate queue reordering policy to be used for job selection. This
approach is assessed by an extensive resampling-based experimental validation
that uses 7 actual traces.

\subsection{Contributions}

Our main contribution is to investigate two original methods for learning a
good scheduling policy on a per-system basis. In this work, we focus on the
average waiting time of the jobs. Each method aims at choosing the best queue
reordering policy in a fixed search set of semantically diverse options.

The first method, \textbf{full feedback}, continuously uses simulation on data gathered
from the past and current system workload. In order to reflect the
uncertainties in the data, we also evaluate a more realistic ``noisy'' variant in
which simulations are imprecise.

The second method, \textbf{bandit feedback}, is a simulator-less approach based on a
multi-armed bandit algorithm, which derives feedback by measuring system
performance.

Both of these methods are evaluated with a specifically developed simulator,
which is open-sourced\cite{ocst}.

Our main results are as follows: The approach based on the simulator provides
very good results, close to the best policy in the search set. In essence, we
are able to recover in an online fashion existing results that reduce average
waiting times by 11\% to 60\% compared to EASY-FCFS.  Indeed, we show
that the required sample sizes for policy selection in batch schedulers are
small enough to avoid workload resampling in practice. The variant
based on noisy data, which is closer to the actual conditions of uncertainties on
both jobs and platform, behaves similarly. The alternate approach based on
a multi-armed bandit also provides reasonable results at a much lower cost,
reducing the average waiting times by 8\% to 46\% compared to
EASY-FCFS.

Let us emphasize the main results of this study: While the First-Come,
First-Serve approach to minimizing the waiting times is a common default
strategy, it can definitively be outperformed.  We show how to automate selection
of a good alternative policy in two cases (depending on whether a simulator is
available or not), and provide the difference in performance to be expected
between both situations.

\section{Related Work}
\label{sec:rw}

This section presents and discusses the most significant results related to job
scheduling and learning algorithms in this context.  Let us start by recalling
some basics results on heuristics in HPC platforms:

Parallel job scheduling is a old studied theoretical
problem~\cite{Frachtenberg:2009:JSS:1692356,Feitelson:2004:PJS:2128864.2128865}
whose practical ramifications, varying hypotheses, and inherent uncertainty of
the problem applied to the HPC field have driven practitioners to use simple
heuristics (and researchers to study their behavior). The two most popular
heuristics for HPC platforms are EASY~\cite{easy} and
Conservative~\cite{Mu'alem:2001:UPW:380314.380315} backfilling.

While Conservative Backfilling offers many advantages~\cite{bfchar}, it has a
significant computational overhead and is more complex, which mainly explains
why many of the machines of the top500 ranking~\cite{top500} rather use a
variant of EASY-Backfilling instead. There is a large body of work seeking to
improve/tune EASY. Indeed, while the basic mechanism is used by some actual
resource and job management software (most notably SLURM~\cite{SLURMdocSCHED}),
this is seldom done without fine tuning by system administrators.

The original EASY mechanism refers to a First-Come-First-Serve basis.  Several
works explore how to tune EASY by reordering waiting and/or backfilling
queues~\cite{Tsafrir_easypp_2005}, sometimes even in a randomized
manner~\cite{1592720}, as well as some implementations~\cite{Jackson2001}.
However, as successful as they may be, these works do not address the
dependency of scheduling metrics on the workload~\cite{variability}.
These studies most often report \textit{post-hoc} performance since they
compare algorithms after the workload is known.

The dynP scheduler~\cite{streit_selftuning_2002} suggests a systematic method
to tuning these queues. Although it requires simulated scheduling runs at
decision time and therefore, making it much more than the natural execution of
EASY.

\subsection{Data-aware Resource Management.}

There was a recent focus on leveraging the huge amount of data available in
large scale computing platforms to improve system performance. Some
works use collaborative filtering to co-locate tasks in clouds by estimating
application interference~\cite{7516031}. Some others are closer to the
application level. For instance, ~\cite{fmodeling} uses binary classification
to distinguish benign memory faults from application errors so as to execute
recovery algorithms.

Several works use similar techniques in the context of HPC, in
particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures~\cite{Nissimov2008}.  However, these works
do not address the duality between the cumulative and maximal scheduling costs,
as mentioned in~\cite{learningruntimes}. While these previous works intend to
estimate uncertain parameters, we consider here a more pragmatic approach,
that directly learns a good scheduling policy from a given policy space.

Existing work~\cite{jsspp17} takes this approach offlines, by
splitting workload data in a training/testing fashion. This work studies
EASY-Backfilling and shows that while the relative performance of some well-known
priority orders for starting jobs differs between workloads, it is relatively
stable over time for a given workload.

\subsection{Multi-Armed Bandits.}

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards\cite{bubnow}. At every round, an action (represented by an arm
of the bandit) must be chosen from a fixed set and the corresponding reward is
observed. The goal of a MAB algorithm is to maximize the total reward obtained
in a successive number of rounds. There exist multiple works addressing this
problem under a variety of constraints, The two most popular settings are the
original stochastic  case and the adversarial case. See~\cite{thompson} for the
original work on the stochastic case and~\cite{Auer2002} for the ''Upper
Confidence Bound`` family of algorithms. To the best of our knowledge, the work
of~\cite{Banos} is the earliest for the adversarial case; see~\cite{nonstoch}
for the ``Exponential-weight'' family of algorithms. We refer to~\cite{bubnow}
for a comprehensive review of the field. While these algorithms bound the
cumulative difference in loss to the best arm (the regret), they have
functional constraints such as the fact that the rewards should be contained in
a range. The simple heuristic called \textit{Epsilon-Greedy} introduced
in~\cite{Auer2002} does not have this requirement.  This heuristic will be used
in this paper and is described in Subsection~\ref{sub:bandit}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Problem Setting}
\label{sec:problem_setting}

%This section presents the systems under study and the scheduling problem.  It
%first introduces the EASY-Backfilling heuristic and gives \todo{fix} the
%problem statement.

\subsection{System Description}
\label{sub:sysdesc}

The crucial part of batch scheduling software is the scheduling algorithm that
determines where and when the submitted jobs are executed. The process is as
follows: jobs are submitted by end-users and queued until the scheduler selects
one of them for running. Each job has a provided bound on the execution time
and some resource requirements (number and type of processing units). Then, the
Resource and Job Management System (RJMS) drives the search for the resources required to execute this job. Finally,
the tasks of the job are assigned to the chosen nodes.

In the classical case, the management software needs to execute a set of
concurrent parallel jobs with rigid (known and fixed) resource requirements on
a HPC platform represented by a pool of $m$ identical resources. This is an
on-line problem since the jobs are submitted over time and their
characteristics are only known when they are released.  Below is the brief
description of the characteristics of job
$j$:

\begin{itemize}

   \item Submission date $r_j$ (also called \textit{release date})

   \item Resource requirement $q_j$ (number of processors)

   \item Actual running time $p_j$ (sometimes called \textit{processing time})

   \item Requested running time $\widetilde{p_j}$ (sometimes called
     \textit{walltime}), which is an upper bound of $p_j$.

\end{itemize}

The resource requirement $q_j$ of job $j$ is known when the job is submitted at
time $r_j$, while the requested running time $\widetilde{p_j}$ is given by the
user as an estimate. Its actual value $p_j$ is only known \textit{a posteriori}
when the job really completes.  Moreover, the users have incentive to
over-estimate the actual values, since jobs may be ``killed'' if they surpass
the provided value.

\subsection{Brief Description of EASY Backfilling}
\label{sub:easy}

The selection of the job to run is performed according to a scheduling policy,
which determines the order the jobs are executed. EASY-Backfilling is
the most widely used policy due to its simple and robust implementation and
known benefits such as high system utilization~\cite{easy}. This strategy has
no worst case guarantee beyond the absence of starvation (i.e.\ every job will
be scheduled at some moment).

EASY-FCFS uses a job queue to select and backfill jobs.  At any
time a scheduling decision is required (i.e.\ job submission or termination),
the scheduler goes through the job queue in First-Come,First-Serve (FCFS) order
and starts jobs until it finds a job that can not be started right away. Then,
it makes a reservation for this job at the earliest predictable time and
starts \textit{backfilling} the job queue in FCFS order using any job that
does not delay the unique reservation.

\subsection{Scheduling Objective}
\label{sub:scheduling_objectives}

A system administrator may use one or multiple cost metric(s). Our study of
scheduling performance relies on the waiting times of the jobs, which is one of
the more commonly used objectives. The waiting time of a job is

     \begin{equation}
       \textbf{Wait}_j =  start_j-r_j
     \end{equation}

where $start_j$ is the starting time of a job.
Like other cost metrics, the waiting time is usually considered in its
\textit{cumulative} version, which means that one seeks to minimize the average
waiting time. It is worth noting that other metrics such as the maximum waiting
time of all the jobs are also worthy of interest. Unfortunately, this criteria
is antagonistic in nature with the average waiting time. Section~\ref{sub:th}
will outline our approach to address the bi-objective aspect of this problem.

\section{Tuning EASY by Reordering and Thresholding the Job Queue}
\label{sec:framework}

This section presents two mechanisms for safely tuning the EASY-Backfilling:
job queue reordering and job thresholding. Together, these two building blocks
make a robust framework for tuning EASY.

\subsection{Reordering}
\label{subsec:policies}

The EASY heuristic uses a job queue to select and backfill jobs. While this job
queue is ordered in FCFS order in the original heuristic, it is possible to
reorder it at will. We choose on a search space of 10 reordering policies.

\begin{enumerate}
  \item FCFS: First-Come First-Serve, which is the most commonly used policy~\cite{easy}.
  \item LCFS: Last-Come First-Serve.
  \item SPF: Smallest estimated Processing time $\widetilde{p_{j}}$ First ~\cite{bfchar}.
  \item LPF: Longest estimated Processing time First.
  \item LQF: Largest resource requirement $q_j$ First.
  \item SQF: Smallest resource requirement First.
  \item LEXP: Largest Expansion Factor First~\cite{bfchar}, where the expansion
    factor is defined as follows:
  \begin{equation} \frac{wait_j + \widetilde{p_j}}{\widetilde{p_j}} \end{equation}
    where $wait_j$ is the time job $j$ has been waiting until the time at which the decision is taken.
  \item SEXP: Smallest Expansion Factor First
  \item LRF: Largest Ratio $\frac{\widetilde{p_j}}{q_j}$ First
  \item SRF: Smallest Ratio First
  \item LAF: Largest Area $ \widetilde{p_j} \times q_j$ First
  \item SAF: Smallest Area First
\end{enumerate}

This search space is designed with the goal of being as semantically diverse as
possible without making any judgement on which policy should perform well in
practice. It does include most policies from related works in this area that we
are aware of. In the following, we denote these policies by $P_i$ with $i = 1
\ldots 12$.

\subsection{Thresholding}
\label{sub:th}

Reordering the job queue means losing the no-starvation guarantee and some
individual jobs therefore can wait an undue amount of time. It is possible to
introduce a thresholding mechanism to prevent this behavior: When a
queued job's waiting time exceeds a fixed threshold $\Theta$, it is
at the head of the queue. We denote by EASY($P,\Theta$) the scheduling
policy that starts and backfill jobs according to the (thresholded) reordering
policy $P$. For the sake of completeness, Algorithm~\ref{alg:EASY} describes
the EASY($P,\Theta$) heuristic.

\begin{algorithm}[]
  \caption{EASY($P,\Theta$) policy}
  \begin{algorithmic}[1]
    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
    \renewcommand{\algorithmicensure}{\textbf{Output:}}
    \REQUIRE Queue $Q$ of waiting jobs.
    \ENSURE None (calls to $Start()$)
    \STATE Sort $Q$ according to $P$
    \STATE Move all jobs of $Q$ for which $wait_j > \Theta$ ahead
    of the queue (breaking ties in FCFS order).
    \\ \textit{//Starting jobs until the machine is full}
    \FOR {job $j$ in Q}
    \IF {$j$ can be started given the current system use}
    \STATE Pop $j$ from Q
    \STATE $Start(j)$
    \ELSE
    \STATE Reserve $j$ at the earliest
    time possible according to the estimated running times
    of the currently running jobs.
    \\ \textit{//Backfilling jobs}
    \FOR {job $j'$ in $Q\setminus\{j\}$}
    \IF {$j'$ can be started without delaying $j$'s reservation.}
    \STATE Pop $j'$ from $Q$
    \STATE $Start(j')$
    \ENDIF
    \ENDFOR
    \STATE \textbf{break}
    \ENDIF
    \ENDFOR
  \end{algorithmic}
  \label{alg:EASY}
\end{algorithm}

\section{Online Tuning}
\label{sec:online}

\begin{figure*}[]
  \centering
  \includegraphics[scale=0.6]{figures/variability.pdf}

  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace
  (see Subsection~\ref{sub:traces}) for 7 different policies with $\Theta = 0$. The policy set is
  reduced as not to obstruct the figure with too many colors.}

  \label{fig:variability}
\end{figure*}

Here, we present our policy selection strategies. We will
refer to the period during which a selected policy is applied as the
\textit{policy period} and will denote the length of this period as $\Delta$.
The time interval is thus divided into periods of equal lengths ($\Delta$):
$\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy
period. A new policy is selected at the beginning of each period and applied
during the whole period.

We further assume that there is a certain regularity among periods, i.e.\ that
the distribution of the job submission process does not radically differ
between consecutive periods. This assumption is validated in the study
presented in~\cite{jsspp17}. It further entails that the behavior of a policy
during the previous periods reflects its behavior on the current one, so that
the selection of a policy can be based on its past behavior. However, there may
be some variability between different periods for certain cost
metrics~\cite{feitelson2001metrics}. This is illustrated in
Figure~\ref{fig:variability} which displays the average waiting time of various
policies for the KTH-SP2 trace (see~\ref{sub:traces} for a description of the
workloads) using weekly periods encompassing one year. As one can note, the
average waiting time varies a lot from one week to the other, for all 7
policies considered. This indicates that when the cost metric is averaged
over different periods, there is a trade-off to find in between longer periods
that would somehow limit the variability, and shorter ones that yield more
values for the estimation. We will come back to this issue below.

The selection of a policy is reminiscent of reinforcement learning, where
solving a search and inference problem in a (perhaps restrictive) policy space
is easier than in the space of the original problem. It is important to note,
however, that a pure reinforcement learning approach is difficult to develop in
our context. Indeed, while we have outlined a reduced action space, the state
space to consider is infinite. This is not prohibitive \textit{per se}, as
modern methods~\cite{aprl} can bypass dimensionality issues via function
approximation. However, these methods rely on large amounts of data, and we are
studying online methods\footnote{The reinforcement learning literature uses the
term ``\textit{on-policy}'' for these methods.}. We rely here on simpler, yet we
believe more effective, strategies to solve this problem.  These strategies are
applied online and rely on exact simulation, noisy simulation and
$\epsilon$-greedy bandit exploration.

\subsection{Policy Selection with Exact Simulation}
\label{sub:feedback}

Several simulators have been developed for ``playing'' reordering policies on a
given set of jobs. We rely in this study on a lightweight simulator (see
Section~\ref{sec:experiments}) that can efficiently simulate different
policies. Such simulators are interesting inasmuch as they provide an estimate
of the cost of a given policy on a set of jobs, as described below.

Let $l(\Delta_t)$ denote the number of jobs \textit{submitted} during the
period $\Delta_t$ ($0 \le t $), and $P_i$ ($1 \le i \le 12$) one reordering
policy (defined in Section~\ref{subsec:policies}). The cost of policy $P_i$
during period $\Delta_t$ is defined as:
%
\begin{equation} \label{eq:cost-exact} w_{\mbox{exact}}(t;P_i) =
\sum_{j=1}^{l(\Delta_t)} \mbox{Wait}_j^{P_i} \end{equation}
%
where $\mbox{Wait}_j^{P_i}$ denotes the estimate provided by the simulator of
the waiting time for job $j$ according to policy $P_i$. The estimation of the
cost of a policy over all the periods preceding the current period can then be
defined as:
%
\begin{equation} \label{eq:tot-cost-exact} w(\rightarrow T;P_i) =
\sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{exact}}(t;P_i) \end{equation}
%
where $\lambda \in [0,1]$ is a decaying discount factor that can be used to
privilege recent history (\textit{i.e.} recent periods). One then selects the
policy $P$ that minimizes the above cost for the period $\Delta_T$:
%
\begin{equation} \label{eq:select-exact} P = \mbox{argmin}_{P_i, 1 \le i \le
12} w(\rightarrow T;P_i) \end{equation}

The above cost directly corresponds to the cumulative waiting time of the
policy over the preceding periods (more precisely to the cumulative simulated
estimate of the waiting time of the policy over the preceding periods) so that
the length of the period has little impact here. The only bias comes from the
boundary states of the simulation. Indeed, we run simulations from an empty
system and wait for the system to be empty when job submissions cease at the
end of the submission period.

Furthermore, in the context of this study and as detailed in
Section~\ref{sec:experiments}, we rely on averages over several execution
traces in order to obtain reliable estimate of the behavior of different
policies and policy selection strategies. Such traces are typically obtained by
simulation. Using the same simulator for generating the traces and estimating
the cost as defined in Eqs \eqref{eq:cost-exact} and \eqref{eq:tot-cost-exact}
would however be too optimistic and would represent an upper bound on what can
be achieved by a selection strategy based on simulation. In order to have a
more realistic estimate of the behavior of a selection strategy based on
simulation, we introduce noise in the simulator, as described below.

\subsection{Policy Selection with Noisy Simulation}
\label{sub:noisy}

In order to simulate how the simulation strategy for selecting policies would
work on non simulated traces, we randomly introduce noise in the estimate of
the policy cost defined by Eq.~\eqref{eq:cost-exact} by rescaling it, either
down or up:
%
\begin{equation} \label{eq:cost-noisy} w_{\mbox{noisy}}(t;P_i) =
 \rho \sum_{j=1}^{l(\Delta_t)} \, \mbox{Wait}_j^{P_i} \end{equation}
%
where $\rho$ is uniformly sampled in the interval $[0.80,1.20]$, thus adding a
+/- 20\% noise on the waiting time estimated by the simulator. The overall cost
is then defined in the same way as before, leading to:
%
\begin{equation} \label{eq:tot-cost-noisy} w(\rightarrow T;P_i) =
\sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i) \end{equation}
%
As before (see Eq.~\ref{eq:select-exact}), the policy that minimizes the above
cost is selected for the period $\Delta_T$.

Instead of introducing noise in the output of the simulator when selecting the
policy, one could have introduced it when collecting the execution traces. Our choice
to add the noise after the simulation step is motivated by simplicity. As we
will see in Section~\ref{sec:experiments}, there is little difference between
the two selection methods, which is an important result of our study.

Lastly, as before, the length of the period has little impact here as the cost
metric corresponds to the cumulative simulated waiting time for each policy.

\subsection{Policy Selection with $\epsilon$-greedy Exploration}
\label{sub:bandit}

The previous strategies estimate the cost of all policies over all previous
periods; the best policy according to this cost is then selected for the
current period, the costs of all policies being updated for the next period.
This is interesting as one maintains a complete knowledge of all policies over
time. However, this requires computing many estimates at each period (as many
as there are policies), which can be time consuming or cumbersome even with
lightweight simulators.

Thus, we explore here a more efficient strategy that dispenses with estimating
the cost of all policies at a given time and only updates the cost of the
policy that is currently being used. This strategy makes use of the
$\epsilon$-greedy exploration method, standard in reinforcement learning
\cite{Watkins:1989} and bandit problems~\cite{Auer2002}, to trade-off
exploitation and exploration, and relies, as before, on past estimates of the
cost to select the best policy in the exploitation mode.

Let $l(\Delta_t)$ denote this time the number of jobs \textit{finished} during
the period $\Delta_t$. We define the cost of policy $P_i$ during period
$\Delta_t$ as:
%
\[ w_{\epsilon}(t;P_i) = \left\{ \begin{array}{ll} \sum_{j=1}^{l(\Delta_t)}
  \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta_t
  \nonumber\\ 0 & \mbox{otherwise} \end{array} \right.  \]
%
As one can note, the actual waiting time is used here, so that one dispenses
with the use of a simulator for efficiency considerations. This leads to a
faster and easier estimate of the cumulative waiting time, however only for the
policy that is used during $\Delta_t$. Note that there is as previously a bias
due to boundary effects in the measurement method, as we do include jobs that
had been started before $\Delta_t$, as well as disregard jobs that were
submitted during $\Delta_t$ but finish later.

%This also explains the difference
%between the definitions of $l(\Delta_t)$: in the previous selection strategies,
%one relies on a simulated estimate so that jobs need not be finished during
%$\Delta_t$, but only submitted; in contrast, to know their actual waiting time,
%jobs need be finished during $\Delta_t$.

The cost over all previous period of a policy $P_i$ is then defined as:
%
\begin{equation} \label{eq:tot-cost-greedy} w(\rightarrow T;P_i) =
\frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)} \sum_{t=0}^{T-1}
\mathbbm{1}(P_i,t) \lambda^{T-1-t} w_{\epsilon}(t;P_i) \end{equation}
%
where $\mathbbm{1}(P_i,t) = 1$ if $P_i$ is the policy used during the period
$\Delta_t$ and $0$ otherwise. The normalization by $\sum_{t=0}^{T-1}
\mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary to ensure that policies
remain comparable over time. Not normalizing would favor the policy that was
first selected, even if this choice was random. This normalization however
entails that the size of the policy period is important: it should not be too
large so as to avoid relying on too few points for estimating the cost, and it
should not be too small either to avoid extreme variations between periods. As
explained in Section~\ref{sec:experiments}, we rely in this study on periods of
one day and one week.

The selection of the policy to be used during the current time period
$\Delta_T$ is then based on the standard $\epsilon$-greedy exploration
strategy:

%
\begin{itemize}
\item With probability $(1-\epsilon)$, select the policy $P$ that minimizes the cost over previous time periods (exploitation mode):
%
\begin{equation}
\label{eq:select-greedy}
P = \mbox{argmin}_{P_i, 1 \le i \le 12} w(\rightarrow T;P_i)
\end{equation}
%
\item With probability $\epsilon$, select a policy $P_i, \, 1 \le i \le 12$, at random (exploration mode).
\end{itemize}

Several studies study the decay of $\epsilon$ over time, the exploration being
less important once the estimates for the different policies are reliable
\cite{Tokic:2010}. This strategy has however not been beneficial in our case.
We believe this is due to the properties of the traces we use, as explained in
Section~\ref{sec:experiments}. The $\epsilon$-greedy strategy is known to be
outperformed in terms of regret by other classical algorithms such as the UCB
or EXP family of algorithms. See~\cite{bubnow} for a precise definition and
thorough treatment of this subject. These algorithms however all rely on
pre-normalized rewards. Here, the scale of $w_{\epsilon}(t;P_i)$ is not known
in advance. While the usual doubling tricks can be devised to go around
sequential issues of scale with good asymptotical performance, they ruin
performance in practice. Accordingly, we only perform experiments using the
simple $\epsilon$-greedy method.

\section{Experiments}
\label{sec:experiments}

This section compares the different approaches to policy selection via a
comprehensive experimental validation.  Subsection~\ref{sub:protocol} outlines
our experimental protocol and Subsection~\ref{sub:results} contains the
experimental results and their discussion.

\subsection{Experimental Protocol}
\label{sub:protocol}

The evaluation of computer systems performance is a complex
task. Here, the two main difficulties are choosing a simulation
model and taking into account the variability in the average waiting times.
This section presents our approach, which is based on lightweight simulation
and trace resampling.

\subsubsection{Traces}
\label{sub:traces}

\begin{table}[]
  \centering
  \ra{1.3}
  \caption{Workload logs used in the simulations.}
  \label{tab:logs}
  \begin{tabular}{@{}lrrrr@{}}
    \hline
    Name          & Year & \# MaxProcs & \# Jobs & Duration\\
    \hline
    KTH-SP2       & 1996 & 100         & 28k     & 11 Months\\
    CTC-SP2       & 1996 & 338         & 77k     & 11 Months\\
    SDSC-SP2      & 2000 & 128         & 59k     & 24 Months\\
    SDSC-BLUE     & 2003 & 1,152       & 243k    & 32 Months\\
    ANL-Intrepid  & 2009 & 163,840     & 68k     & 9  Months\\
    CEA-Curie     & 2012 & 80,640      & 312k    & 3  Months\\
    Unilu-Gaia    & 2014 & 2,004       & 51k     & 4  Months\\
    \hline
  \end{tabular}
\end{table}

The experimental validation performed in this work uses a mixture of small and
large systems from different periods. Table~\ref{tab:logs} presents the 7 logs,
which can all be obtained from the Parallel Workload
Archive~\cite{Feitelson20142967}. We use the ``cleaned''\cite{pwa-clean} version
of the logs as per the archive. We impose an additional filtering step to the
workloads\footnote{This filtering step is available in the reproducible
workflow\cite{repro} as shell script
\lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter|} in order to
clean erroneous data:

\begin{enumerate}

  \item If the number of allocated or requested cores of a job  exceeds the
    size of the machine, we remove the job;

  \item If the number of allocated or requested cores is negative, we use the
    available positive value as the request. If both are negative, we remove
    the job;

  \item If the runtime or submission time is negative, we remove the job.

\end{enumerate}

\subsubsection{Simulator}

The choice of a simulator is a critical part of experimental validation that
raises a trade-off between precision and runtime. High precision can be
obtained by carefully modeling the platform and its network topology, or
extracting information from the jobs from their sources or post-mortem logs.
This is the approach used by high-fidelity batch scheduling simulators such as
BatSim~\cite{batsim}. In the present work, the focus is on studying the
EASY-Backfilling mechanism in itself, without addressing the allocation
problem. Moreover, the experimental protocol used here requires many
simulation runs. Therefore, we set the precision/runtime trade-off at the
point which minimizes simulation runtimes. We discard all topological
information relative to the platform and use the job processing times of the
original workloads. In this setup, the processors are considered to be
undistinguishable from each other, and jobs can be discontinuously mapped to
any available processor on the system. During this work, we developed a
lightweight simulator\cite{ocst} and now release it to the community along with
this article. See the reproducibility paragraph below for more information.
With this simulator, we are able to replay EASY-FCFS on the CEA-CURIE trace in
33 seconds with a machine equipped with an Intel(R) Core(TM) i5-4310U CPU @
2.00GHz. As a comparison, the Batsim simulator with network communication
modeling disabled and no resource contiguity takes more than 7 hours to replay
the same trace on the same hardware. The complete set of simulations presented
below is executed on a single Dell PowerEdge R730 in 22 hours, including input
preparation and statistical analysis code.

\subsubsection{Resampling}
\label{ssub:resampling}

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics~\cite{jsm}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. In our experience, randomly
shuffling months in a trace at random is enough to generate contradictory
conclusions about which algorithm is better. This is a known
problem~\cite{flurries} that motivated decades of research in workload
modeling~\cite{feitbook}. A recent approach~\cite{feitresampling} suggests to
resample data directly from the original traces. In this work, we use a simple
implementation of these ideas.  More precisely, we want to sample the job
submission process, which is achieved in the following manner: (a) For every
week in a resampled trace, we iterate over every existing user in the original
trace; (b) For each of these users, we choose a week of the original trace
uniformly at random; (c) The jobs from this user in this original week are
chosen to be present in the week being resampled.  This simple mechanism
enables us to preserve the dependency within a week.  Although the system state
may not be independent from one week to an other, we rely on existing
work~\cite{jsspp17} and assume for these experiments that the system is in a
stationary state. While we do provide guidelines as how to incorporate tracking
variant of our algorithms in Section~\ref{sec:online} by introducing a decaying
discount factor $\lambda$, this will not be studied in the experiments where 
$\lambda=1$, and is left for future research.

%Thus,
%computing an average over the sampled traces at a given time can be related to
%a time average of an underlying Markov process, which converges to the true
%quantity thanks to the ergodic theorem.  Moreover, the job submission process
%has no long range correlation making it look like an independent and
%identically distributed process when considering sufficiently time spaced
%weeks.

\subsubsection{Experimental Validation}

For each of the 7 workloads
considered, we use the original data to resample 60 traces, each with a length
of two years. For each resampled trace, we run simulations using the
EASY($P,\Theta$) scheduling heuristic. We fix the value of the job waiting time
threshold $\Theta$ at 40 hours. The choice of this default
stems from the analysis developed in ~\cite{jsspp17}. Essentially, this value is high
enough for queue reorderings to be leveraged while being low enough to provide
the functional requirement of preventing starvation. We simulate
EASY($P,\Theta=40h$) using the following strategies:

\begin{itemize}

  \item All fixed $P_i$ for $1 \leq i \leq 12$ from the search set defined in
    Subsection ~\ref{subsec:policies}.  This is done for two reasons. First, it allows to
    see which policy works well for which platform. Second, it allows to
    compare how well strategies used in this work with the best
    \textit{a-posteriori} known policy;

  \item For all the above strategies based on exact simulation (referred to as
    ``Simulated feedback'' in the remainder), on noisy simulation (referred to as
    ``Noisy Simulated Feedback'') and on $\epsilon$-greedy exploration (referred to as
    ``Bandit Feedback''), we experiment with two policy periods: $\Delta \in
    \left\{ 1 \text{day},1 \text{week} \right\}$. This choice is a consequence
    of the natural rhythm of human activity: Daily or weekly simulation periods
    seem to ease comparing of performance between two periods.
    Indeed, arbitrary period values would needlessly increase the variability
    of the cost metric. For the bandit feedback strategy, $\epsilon$ is set to $0.5$ and
    $\lambda$ to $1$. This means that we do not use any decay in the estimation
    of the overall cost. Choosing to eliminate decay in this experimental
    validation is natural since our resampled data is in a stationary regime
    \textit{by construction}, as explained in~\ref{ssub:resampling}. The choice
    of $\epsilon=0.5$ is a reasonable default;

  \item Lastly, an additional ``Random'' strategy is used as a reference point,
    with, as before, $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$.
    This strategy corresponds to choosing $P_i$ uniformly at random in
    $\left\{P_1 , \cdots, P_{12} \right\}$ during every period $\Delta_t$.

\end{itemize}

The experimental validation uses of 7 traces $\times$ 60 resamples $\times$
(12+2+2+2) algorithms, which correspond to 12600 scheduling simulations spanning
two years.  This means a total simulated timespan of 25200 years. Moreover,
both the full-feedback and noisy feedback strategies will also be calling
simulations as part of their procedure. Indeed, these strategies re-simulate
the system 12 times at every period $\Delta$, which bumps the total
effective simulation timespan to 92400 years. This motivates the use of a
lightweight simulator in the context of this research.

\subsection{Replicability}

We publish a declarative workflow that runs simulations and generates all the
figures from this paper. Any question regarding replication can be freely
directed to the authors via e-mail.

There are various approaches for ensuring computational experiments are replicable,
among which distributing complete operating system images, containers, or
packaging software. As our experiments are CPU bound, we decide to opt for the
software packaging approach\ref{nix}. This allows to replicate the experiments on
clusters where virtualization is not available or kernels are too old for
containers and decreases simulation runtime.

All the dependencies of our
experiments (including our own code, dependencies for data processing and
visualization and workflow engine) are automatically managed by Nix up to the
actual execution of the code.
The Nix package manager is available at \url{https://nixos.org/nix}.

The archive containing the experimental workflow can be obtained at:
\begin{lstlisting}
https://github.com/freuk/obps
\end{lstlisting}

All the code associated with this work is released under the~\cite{isc}
license.
Running the experiments can be done
on any platform equipped with the Nix package manager by running at the root of
the extracted archive:

\begin{lstlisting}
nix-build -A banditSelection
\end{lstlisting}

The experiments in this paper can run in less than 24 hours on a
moderately parallel host. More precisely, we had access to a Dell PowerEdge
R730 equipped with a total of 56 threads @2.4GHz each and 757 Gigabites of RAM. For
this reason, the experimental workflow was designed to be executed on
a single host.

All data from this paper were obtained from the Swf Parallel Workloads
Archive~\cite{Feitelson20142967}.

 All experiments are tied together using zymake~\cite{zymake}, a minimalistic
 workflow system designed for computational experiments\footnote{The zymake system
 is also packaged by our nix expressions.}. This system is
 analogous to a traditional build system with added workflow capabilities. The
 entire workflow that generates this article from the input data is contained
 in a single \lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile| to be
 found at the root of the main archive. This workflow is composed of the
 following steps:

 \begin{itemize}
   \item Data. Data extraction from archives, filtering, resampling.
     this is principally using shell scripts for data manipulation (see file
     \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong_filter| for the
     filtering steps used) and ocaml code that implements the resampling method
     described in~\ref{ssub:resampling}.

   \item Simulation. This step runs the lightweight ocaml backfilling
     simulator specially written for this work. this simulator is made
     available under the isc~\cite{isc} license as a persistent zenodo
     archive at~\cite{ocst}.

   \item Analysis. This step runs R code that generates the figures
     presented in this paper.
 \end{itemize}

All resulting figures will be situated in the simlinked folder named
\lstinline[basicstyle=\ttfamily\color{blue}]|result| positionned at the root of
the archive.

\subsection{Results}
\label{sub:results}

\subsubsection{Analysis of Basic Policies}

\begin{table*}[]
  \centering
  \ra{1.3}

  \caption{Average total waiting time diminution of fixed policies with respect
  to EASY(FCFS). The precise definition of the value reported is provided in
Eq~\eqref{eq:cum}}

  %DELTA WEEK

  \label{tab:full}
  \begin{tabular}{@{}lrrrrrrrrrrr@{}}
    \hline
    Trace        & LCFS  & LPF            & SPF   & LQF            & SQF   & SEXP  & LEXP           & LRF            & SRF   & LAF            & SAF   \\
    \hline
    KTH-SP2      & -13\% & \textbf{-16\%} & +5\%  & \textbf{-16\%} & +3\%  & -8\%  & -15\%          & -8\%           & -13\% & -12\%          & +15\% \\
    CTC-SP2      & -40\% & -19\%          & -14\% & \textbf{-47\%} & +15\% & -36\% & -19\%          & -38\%          & -12\% & -44\%          & +22\% \\
    SDSC-SP2     & -8\%  & \textbf{-15\%} & +4\%  & -8\%           & +4\%  & -3\%  & \textbf{-15\%} & -3\%           & -10\% & -11\%          & +18\% \\
    SDSC-BLUE    & -26\% & -29\%          & 0\%   & -29\%          & +17\% & -17\% & -26\%          & -15\%          & -15\% & \textbf{-32\%} & +23\% \\
    ANL-Intrepid & -25\% & -9\%           & -3\%  & \textbf{-28\%} & -2\%  & -21\% & -24\%          & -21\%          & -10\% & -19\%          & +5\%  \\
    CEA-Curie    & -46\% & -23\%          & -45\% & \textbf{-57\%} & -19\% & -51\% & -23\%          & \textbf{-57\%} & -24\% & -40\%          & -17\% \\
    Unilu-Gaia   & -58\% & -33\%          & -10\% & \textbf{-62\%} & +16\% & -49\% & -27\%          & -56\%          & -16\% & -62\%          & +20\% \\
    \hline
  \end{tabular}
\end{table*}

We begin by drawing the attention of the reader to Table~\ref{tab:full} that
reports the average total waiting time performance improvement over
EASY(FCFS,40h) of all the fixed policies defined in ~\ref{subsec:policies}. Let
us denote the resampled traces of a given workload by $T_k$, where k is the
index of the resampled trace ($1 \le k \le 60$). Let the number of jobs in the
resampled trace $T_k$ be $l(T_k)$. Let $\bar{w}(T_k,P)$ be the cumulative
waiting time obtained by a simulation run for a strategy $P$:

\begin{equation}
  \bar{w}(T_k,P) = \sum_{j=1}^{l(T_k)} \text{Wait}_j^{P}
\end{equation}

So that the values appearing in Table~\ref{tab:full} are exactly:

\begin{equation} \label{eq:cum}
  \frac{\sum_{k=1}^{k=60}\bar{w}(T_k,\text{EASY}(P_i,40h))-
  \bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}{\sum_{k=1}^{k=60}\bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}
\end{equation}

The best values are outlined in bold in the table. These values give the best
attainable performance in the search set defined in~\ref{subsec:policies}.
Note that there seems to be some regularity; for instance the SAF policy is remarkably
inefficient and the LQF policy works very well. However there are traces for which LQF is not optimal. Moreover, these
are~\textit{a-posteriori} results, which means that we only know
\textit{post-factum} that this policy would have been better.

\begin{figure}[]
  \centering
  \includegraphics[scale=0.6]{figures/full-UniLu-Ga.pdf}
  \caption{Evolution of the average cumulative waiting time improvement
    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 12$ on the
    UniLu-Gaia trace. The average is obtained by resampling the original trace
    60 times. The dashed lines represent the 10th and 90th percentiles of the
  values across this resampling.}
  \label{fig:all}
\end{figure}

In the present work, we are studying adaptive policies that obtain a good
performance on any given system by selecting one of these policies.
Figure~\ref{fig:all} reports the evolution of the average cumulative waiting
time for the UniLu-Gaia trace. This figure shows how the average improvement
in cumulative waiting time evolves as a function of time.  In other words, this
is the average curve obtained if one were to plot the cumulative sum of job
waiting times for every resampled trace, incrementing the sum when a job
finishes. Accordingly, the final values attained by every curve correspond
exactly to the values reported in Table~\ref{tab:full}. The dashed values
represent the 10th and 90th percentile of the values.

Observe that while the 10th and 90th percentiles of the improvement with
respect to EASY-FCFS are large, the values are still contained enough to make
definite statements about how these policies compare to the FCFS baseline.
This figure does not report the variability of the policies against one another,
and does not allow to compare them beyond their average performance.
The fact that the size of the percentile band is roughly half
the best attainable improvement was our original motivation for this work, as
one can expect to be able to distinguish between these policies in an
online manner.

\subsubsection{Analysis of Policy Selection Strategies}

\begin{figure*}[]
  \centering
  \includegraphics[scale=0.6]{figures/CTC-SP2.pdf}
  \includegraphics[scale=0.6]{figures/KTH-SP2.pdf}\\
  \includegraphics[scale=0.6]{figures/CEA-Curi.pdf}
  \includegraphics[scale=0.6]{figures/SDSC-BLU.pdf}\\
  \includegraphics[scale=0.6]{figures/SDSC-SP2.pdf}
  \includegraphics[scale=0.6]{figures/ANL-Intr.pdf}

  \caption{Evolution of the average cumulative waiting time improvement
    compared to EASY-FCFS of the Simulated Feedback, Noisy Simulated Feedback and Bandit Feedback
    policies. The average is obtained by resampling the original trace 60
    times. The dashed lines represent the 10th and 90th percentiles of the
  values across this resampling. Each figure is a different trace. This figure
is followed-up in figure~\ref{fig:follow} for the remaining UniLu-Gaia trace.}

  \label{fig:small}
\end{figure*}

\begin{figure}[]
  \centering
  \includegraphics[scale=0.6]{figures/UniLu-Ga.pdf}

  \caption{Follow-up from figure~\ref{fig:small}, plot for the UniLu-Gaia log.}

  \label{fig:follow}
\end{figure}

Figures~\ref{fig:small} and~\ref{fig:follow} report the main experimental
results concerning the strategies used in this paper. These figures show the
average cumulative improvement of over EASY(FCFS,40h) of the three strategies
we use in the same format as Figure~\ref{fig:all}. There are five
strategies displayed on the graph:

\begin{itemize}
  \item Random $P \in P_i, 1 \le k \le 60$. ($\Delta=$1 week)
  \item Bandit Feedback ($\Delta=$1 week)
  \item Bandit Feedback ($\Delta=$1 day)
  \item Simulated Feedback ($\Delta=$1 week)
  \item Noisy Simulated Feedback ($\Delta=$1 week)
\end{itemize}

\begin{table*}[]
  \centering
  \ra{1.3}
  \caption{Average Cumulative waiting time improvement of the policies used with respect to EASY(FCFS).}
  \label{tab:strat}
  \begin{tabular}{@{}lrrrrrrrrrrr@{}}
    \hline
    Trace        & Random &       & Full           &                & Noisy &       & Bandit & \\
    \hline
    $\Delta$     & week   & day   & week           & day            & week  & day   & week   & day \\
    KTH-SP2      & -6\%   & -8\%  & \textbf{-12}\% & -\textbf{11}\% & -12\% & -12\% & -7\%   & -10\% \\
    CTC-SP2      & -20\%  & -23\% & \textbf{-44}\% & \textbf{-44}\% & -44\% & -44\% & -26\%  & -31\% \\
    SDSC-SP2     & -4\%   & -7\%  & \textbf{-13}\% & \textbf{-13}\% & -13\% & -13\% & -5\%   & -8\% \\
    SDSC-BLUE    & -12\%  & -14\% & \textbf{-31}\% & \textbf{-31}\% & -30\% & -29\% & -12\%  & -17\% \\
    ANL-Intrepid & -13\%  & -20\% & -22\%          & \textbf{-26}\% & -19\% & -24\% & -13\%  & -20\% \\
    CEA-Curie    & -35\%  & -42\% & \textbf{-53}\% & -47\%          & -53\% & -48\% & -36\%  & -46\% \\
    Unilu-Gaia   & -29\%  & -31\% & -59\%          & \textbf{-60}\% & -58\% & -58\% & -33\%  & -34\% \\
    \hline
  \end{tabular}
\end{table*}

Table~\ref{tab:strat} reports the average total waiting time performance
improvement over EASY(FCFS,40h) of all the strategies used along with the
random variant.  We make the following observations.

\textbf{FCFS vs Random.} Randomly taking a policy in the search set defined in
~\ref{subsec:policies} is better than using the FCFS order. This is due to the
known fact that the FCFS order is not optimal for optimizing the average
waiting time objective, and confirms that the search set is well chosen.  The
value of the period length $\Delta$ has an impact on the performance of the
random policy. Indeed, shorter periods lead to better performance. This is perhaps
due to the use of queuing in the scheduling process:
When a bad policy is used, the job queue builds up. By switching more
frequently from bad to good policies, the system never has time to build up the
job queue. In other words, while bad decisions impact the system through an
accumulation of waiting jobs, there is less contribution to system inertia from
good decisions, as the number of jobs in queue can not be negative. This
interpretation also means that randomized bandit strategies are robust by
design, and needs more subsequent experimenting to be validated.

\textbf{The Simulated Feedback strategy consistently performs almost as well as the
best fixed policy.} The values attained by the exact simulation based strategy almost match
the highlighted values from Table~\ref{tab:full}. This corresponds to the curve
labeled ``Full'' in Figures~\ref{fig:small} and~\ref{fig:follow}. Moreover,
these average cumulative waiting time curves exhibit linearity (as opposed to
being concave), which further shows that the choice of the best policy can be
made rapidly.  Note that the value of $\Delta$ has no impact over the
performance of this strategy.

\textbf{Noisy simulations can be used.} The performance attained by the noisy
variant almost matches that of the exact-simulation based strategy. This indicates
that it is not necessary to have access to a precise simulator. Rather, we may
hope that an imprecise but unbiased reporting of the performance is enough to
rapidly select a reordering policy.

\textbf{Full vs Bandit feedback.} As expected, the bandit strategy evolves
between the random strategy and the full feedback strategy. The bandit-based
strategy always fares better with a shorter period $\Delta$, recovering between
a quarter to the whole gap between the Random and Full Feedback strategies.

\begin{figure}[]
  \centering
  \includegraphics[scale=0.6]{figures/mosaic-UniLu-Ga.pdf}

  \caption{Share of the policies chosen by the Noisy policy ($\Delta=$1 day,
  $\lambda = 1 $) as a function of time
    for the UniLu-Gaia trace. The average choice
  is obtained by resampling the original trace 60 times and aggregating by
  date. The legend excludes policies that are never used by
  the strategy.}

  \label{fig:mosn}
\end{figure}

\textbf{Stability of the policy selection.}
Figure~\ref{fig:mosn} gives insight regarding the behavior of the noisy
feedback policy with $\Delta=$ 1 week. This figure reports the
share of the policies chosen by the strategy on average as a function of time.
This is an area plot: The height of the colored region
represents the value. Moreover, the proportion of times each policy is
chosen by the strategy is aggregated across resampled traces.

While this policy obtains an excellent performance, the convergence to a fixed
policy is slow. This is not harmful \textit{per se} and just means that it is
difficult to distinguish between the best policies. Additionally, we observe that
the two increasingly dominant policies, LAF and LQF, are the best overall policies.
This means that the estimation converges to the correct point.

\begin{figure}[]
  \centering
  \includegraphics[scale=0.6]{figures/mosaicbandit-UniLu-Ga.pdf}

  \caption{Share of the policies chosen by the Bandit Feedback($\Delta=$1 day,
  $\epsilon = 0.5, \lambda = 1 $) policy as a function of time for the UniLu-Gaia
trace. The average choice is obtained by resampling the original trace 60 times
and aggregating by date. The legend excludes policies that are never used by
  the strategy.}

  \label{fig:mosb}
\end{figure}

Figure~\ref{fig:mosb} is the analogue of Figure~\ref{fig:mosn}, this time for
the Bandit Feedback policy with $\Delta$ = 1 week.  Observe that here the choice of policy is much less stable. The convergence is
extremely slow, which is expected given the variability of the reward we give
to the bandit algorithm.

\section{Conclusion}
\label{sec:ccl}

The scheduling of parallel jobs on a given HPC platform is a very hard
optimization problem with many uncertain parameters. Even though these
uncertainties could be reduced, determining efficient strategies remains
difficult.

In this work, we presented a new way of addressing this problem. The idea was
to look directly at the scheduling process instead of trying to change its
parameters.

More precisely, we showed that it is worth learning on the scheduling process
itself. Indeed, reordering the submission queues under EASY-Backfilling leads
to considerable gains in performance.

This approach was used in two methods: a method based on a simulator and a
method based on a multi-armed bandit algorithm. The first one provides very
good results, reducing the average waiting times of the baseline FCFS ordering
policy by 11\% to 60\%. The second one is slightly less efficient with an
improvement factor of 8\% to 46\%. However, the bandit-based method is easier
to use and cheaper to run.

There are two directions to extend this work.  First, the question of
the existence of a more efficient selection strategy for bandit feedback for
these systems is not addressed by this work. Is a simulator really necessary?
%@Denis: a essayer.. 1 mois de travail pour tester?
Second, one can't help but wonder what performance could be achieved by extending
the search space to a wider class of policies. This means studying learning performance using some
function approximator.
%@Denis: c'est ce que font vasily et salah de deux manieres differentes.
%TODO SPLIT IN TWO, MENTION FACTOR ANALYSIS.
As both these directions are data-bound, they will require careful application of
trace resampling techniques to be successful.

\section*{Acknowledgements}

Authors are sorted in alphabetical order. The authors would like to warmly
thank Pierre Neyron for help with computational experiments, as well as Michael
Mercier, Salah Zrigui, Bruno Raffin and Theophile Terraz for their comments and
Arnaud Legrand for his work on reproducibility. We gracefully thank the
contributors of the Parallel Workloads Archive, Victor Hazlewood (SDSC SP2),
Travis Earheart and Nancy Wilkins-Diehr (SDSC Blue), Lars Malinowsky (KTH SP2),
Dan Dwyer and Steve Hotovy (CTC SP2), Joseph Emeras (CEA Curie and UniLu Gaia),
Susan Coghlan, Narayan Desay, Wei Tang (ANL Intrepid), and of course Dror
Feitelson.  The Metacentrum workload log was graciously provided by the Czech
National Grid Infrastructure MetaCentrum. This work has been partially
supported by the LabEx
PERSYVAL-Lab(ANR-11-LABX-0025-01) funded by
the French program Investissement d'avenir.
 Experiments presented in this paper were carried out using the Grid'5000
 testbed. Grid'5000 is supported by a scientific interest group hosted by Inria
 and including CNRS, RENATER and several Universities as well as other
 organizations\footnote{https://www.grid5000.fr}.  Access to the experimental
 machine(s) used in this paper was gracefully granted by research teams from
 LIG\footnote{http://www.liglab.fr} and Inria\footnote{http://www.inria.fr}.

\bibliographystyle{abbrv}
\bibliography{bibliography}

\end{document}

### 2018-01-07 paper/src/paper.tex
modifs

@@
 A multi-armed bandit (MAB) problem is a sequential allocation problem with
-partially observable rewards\cite{bubnow}. At every round, an action (represented by an arm
-of the bandit) must be chosen from a fixed set and the corresponding reward is
-observed. The goal of a MAB algorithm is to maximize the total reward obtained
-in a successive number of rounds. There exist multiple works addressing this
-problem under a variety of constraints, The two most popular settings are the
-original stochastic  case and the adversarial case. See~\cite{thompson} for the
-original work on the stochastic case and~\cite{Auer2002} for the ''Upper
-Confidence Bound`` family of algorithms. To the best of our knowledge, the work
-of~\cite{Banos} is the earliest for the adversarial case; see~\cite{nonstoch}
-for the ``Exponential-weight'' family of algorithms. We refer to~\cite{bubnow}
-for a comprehensive review of the field. While these algorithms bound the
-cumulative difference in loss to the best arm (the regret), they have
-functional constraints such as the fact that the rewards should be contained in
-a range. The simple heuristic called \textit{Epsilon-Greedy} introduced
-in~\cite{Auer2002} does not have this requirement.  This heuristic will be used
-in this paper and is described in Subsection~\ref{sub:bandit}
+partially observable rewards\cite{bubnow}. At every round, an action called an
+"arm" in the litterature must be chosen from a fixed set and the corresponding
+reward is observed. The goal of a MAB algorithm is to maximize the total reward
+obtained in a successive number of rounds. This is essentially achieved using a
+combination of explorative actions, which help estimate the quality of each
+arm, and of exploitative actions, which leverage the current estimates by using
+a good arm. This problem can be addressed using various hypotheses. Most common
+are the (original) stochastic setting and the adversarial setting of regret
+minimization. See~\cite{thompson} for the original work on the stochastic
+setting and~\cite{Auer2002} for the ''Upper Confidence Bound`` family of
+algorithms. To the best of our knowledge, the work of~\cite{Banos} is the
+earliest for the adversarial setting; see~\cite{nonstoch} for the
+``Exponential-weight'' family of algorithms. We refer to~\cite{bubnow} for a
+comprehensive review of the field. While these algorithms bound the cumulative
+difference in loss to the best arm (the regret), they have functional
+constraints such as the fact that the rewards should be contained in a range.
+The simple heuristic called \textit{Epsilon-Greedy} introduced
+in~\cite{Auer2002} does not have this requirement. For this reason, this
+heuristic will be used in this paper and is described in
+Subsection~\ref{sub:bandit}.
 
@@
 queue is ordered in FCFS order in the original heuristic, it is possible to
-reorder it at will. We choose on a search space of 10 reordering policies.
+reorder it at will. We focus on a search space of 10 reordering policies.
 
@@
     \FOR {job $j'$ in $Q\setminus\{j\}$}
-    \IF {$j'$ can be started without delaying $j$'s reservation.}
+    \IF {$j'$ can be started without delaying the reservation on $j$.}
     \STATE Pop $j'$ from $Q$
@@
 
-Let $l(\Delta_t)$ denote this time the number of jobs \textit{finished} during
-the period $\Delta_t$. We define the cost of policy $P_i$ during period
-$\Delta_t$ as:
-%
-\[ w_{\epsilon}(t;P_i) = \left\{ \begin{array}{ll} \sum_{j=1}^{l(\Delta_t)}
-  \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta_t
+Let $l(\Delta^{'}_t)$ denote this time the number of jobs \textit{finished} during
+the period $\Delta^{'}_t$. We define the cost of policy $P_i$ during period
+$\Delta^{'}_t$ as:
+%
+\[ w_{\epsilon}(t;P_i) = \left\{ \begin{array}{ll} \sum_{j=1}^{l(\Delta^{'}_t)}
+  \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta^{'}_t
   \nonumber\\ 0 & \mbox{otherwise} \end{array} \right.  \]
@@
 faster and easier estimate of the cumulative waiting time, however only for the
-policy that is used during $\Delta_t$. Note that there is as previously a bias
+policy that is used during $\Delta^{'}_t$. Note that there is as previously a bias
 due to boundary effects in the measurement method, as we do include jobs that
-had been started before $\Delta_t$, as well as disregard jobs that were
-submitted during $\Delta_t$ but finish later.
+had been started before $\Delta^{'}_t$, as well as disregard jobs that were
+submitted during $\Delta^{'}_t$ but finish later.
 
 %This also explains the difference
-%between the definitions of $l(\Delta_t)$: in the previous selection strategies,
+%between the definitions of $l(\Delta^{'}_t)$: in the previous selection strategies,
 %one relies on a simulated estimate so that jobs need not be finished during
-%$\Delta_t$, but only submitted; in contrast, to know their actual waiting time,
-%jobs need be finished during $\Delta_t$.
+%$\Delta^{'}_t$, but only submitted; in contrast, to know their actual waiting time,
+%jobs need be finished during $\Delta^{'}_t$.
 
@@
 \begin{equation} \label{eq:tot-cost-greedy} w(\rightarrow T;P_i) =
-\frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)} \sum_{t=0}^{T-1}
+  \frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta^{'}_t)} \sum_{t=0}^{T-1}
 \mathbbm{1}(P_i,t) \lambda^{T-1-t} w_{\epsilon}(t;P_i) \end{equation}
@@
 where $\mathbbm{1}(P_i,t) = 1$ if $P_i$ is the policy used during the period
-$\Delta_t$ and $0$ otherwise. The normalization by $\sum_{t=0}^{T-1}
-\mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary to ensure that policies
+$\Delta^{'}_t$ and $0$ otherwise. The normalization by $\sum_{t=0}^{T-1}
+\mathbbm{1}(P_i,t) l(\Delta^{'}_t)$ is necessary to ensure that policies
 remain comparable over time. Not normalizing would favor the policy that was
@@
   \item If the number of allocated or requested cores is negative, we use the
-    available positive value as the request. If both are negative, we remove
-    the job;
-
-  \item If the runtime or submission time is negative, we remove the job.
+    available positive value as the request. If both are negative (no data is
+    available), we remove the job;
+
+  \item If the runtime or submission time is negative (no data is available),
+    we remove the job.
 
@@
 Comparing algorithms in batch scheduling is made especially difficult by the
-fluctuations in commonly used metrics~\cite{jsm}. The variability in
+fluctuations in commonly used metrics~\cite{feitelson1998metrics}. The variability in
 performance is known to outrank the available sample sizes: \textit{It is not
@@
 problem~\cite{flurries} that motivated decades of research in workload
-modeling~\cite{feitbook}. A recent approach~\cite{feitresampling} suggests to
-resample data directly from the original traces. In this work, we use a simple
-implementation of these ideas.  More precisely, we want to sample the job
+modeling~\cite{feitbook}. Workload models have drawbacks in that they usually
+lose the details of the workloads. This is problematic due to the non-linear
+aspect of the job scheduling process. This is discussed
+in~\cite{feitresampling}, an approach which attempts to solve this problem by
+resampling data directly from the original traces. In this work, we use a
+simple implementation of these ideas. More precisely, we want to sample the job
 submission process, which is achieved in the following manner: (a) For every
@@
 uniformly at random; (c) The jobs from this user in this original week are
-chosen to be present in the week being resampled.  This simple mechanism
-enables us to preserve the dependency within a week.  Although the system state
-may not be independent from one week to an other, we rely on existing
+chosen to be present in the week being resampled. This simple mechanism enables
+us to preserve the dependency within a week. Although the system state may not
+be independent from one week to an other, we rely on existing
 work~\cite{jsspp17} and assume for these experiments that the system is in a
@@
 
-We publish a declarative workflow that runs simulations and generates all the
-figures from this paper. Any question regarding replication can be freely
-directed to the authors via e-mail.
+We publish a declarative workflow~\cite{repro} based on our
+simulator~\cite{ocst} that runs experiments and generates all the figures from
+this paper.
 
@@
 packaging software. As our experiments are CPU bound, we decide to opt for the
-software packaging approach\ref{nix}. This allows to replicate the experiments on
+software packaging approach~\cite{nix}. This allows to replicate the experiments on
 clusters where virtualization is not available or kernels are too old for
@@
 
-All the code associated with this work is released under the~\cite{isc}
+All the code associated with this work is released under the ISC~\cite{isc}
 license.
@@
    \item Data. Data extraction from archives, filtering, resampling.
-     this is principally using shell scripts for data manipulation (see file
+     This is principally using shell scripts for data manipulation (see file
      \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong_filter| for the
@@
      simulator specially written for this work. this simulator is made
-     available under the isc~\cite{isc} license as a persistent zenodo
+     available under the ISC~\cite{isc} license as a persistent zenodo
      archive at~\cite{ocst}.
@@
 
-  \caption{Follow-up from figure~\ref{fig:small}, plot for the UniLu-Gaia log.}
+  \caption{Figure~\ref{fig:small} continued, plot for the UniLu-Gaia log.}
 
@@
 results concerning the strategies used in this paper. These figures show the
-average cumulative improvement of over EASY(FCFS,40h) of the three strategies
+average cumulative improvement over EASY(FCFS,40h) of the three strategies
 we use in the same format as Figure~\ref{fig:all}. There are five
@@
   \label{tab:strat}
-  \begin{tabular}{@{}lrrrrrrrrrrr@{}}
+  \begin{tabular}{@{}lrrrrrrrrrrrr@{}}
     \hline
-    Trace        & Random &       & Full           &                & Noisy &       & Bandit & \\
+    Trace        & Best(hindsight) & Random &       & Full           &                & Noisy &       & Bandit & \\
     \hline                
-    $\Delta$     & week   & day   & week           & day            & week  & day   & week   & day \\
-    KTH-SP2      & -6\%   & -8\%  & \textbf{-12}\% & -\textbf{11}\% & -12\% & -12\% & -7\%   & -10\% \\
-    CTC-SP2      & -20\%  & -23\% & \textbf{-44}\% & \textbf{-44}\% & -44\% & -44\% & -26\%  & -31\% \\
-    SDSC-SP2     & -4\%   & -7\%  & \textbf{-13}\% & \textbf{-13}\% & -13\% & -13\% & -5\%   & -8\% \\
-    SDSC-BLUE    & -12\%  & -14\% & \textbf{-31}\% & \textbf{-31}\% & -30\% & -29\% & -12\%  & -17\% \\
-    ANL-Intrepid & -13\%  & -20\% & -22\%          & \textbf{-26}\% & -19\% & -24\% & -13\%  & -20\% \\
-    CEA-Curie    & -35\%  & -42\% & \textbf{-53}\% & -47\%          & -53\% & -48\% & -36\%  & -46\% \\
-    Unilu-Gaia   & -29\%  & -31\% & -59\%          & \textbf{-60}\% & -58\% & -58\% & -33\%  & -34\% \\
+    $\Delta$     &        & week   & day   & week           & day            & week  & day   & week   & day \\
+    KTH-SP2      & -16\%  & -6\%   & -8\%  & \textbf{-12}\% & -\textbf{11}\% & -12\% & -12\% & -7\%   & -10\% \\
+    CTC-SP2      & -47\%  & -20\%  & -23\% & \textbf{-44}\% & \textbf{-44}\% & -44\% & -44\% & -26\%  & -31\% \\
+    SDSC-SP2     & -15\%  & -4\%   & -7\%  & \textbf{-13}\% & \textbf{-13}\% & -13\% & -13\% & -5\%   & -8\% \\
+    SDSC-BLUE    & -32\%  & -12\%  & -14\% & \textbf{-31}\% & \textbf{-31}\% & -30\% & -29\% & -12\%  & -17\% \\
+    ANL-Intrepid & -28\%  & -13\%  & -20\% & -22\%          & \textbf{-26}\% & -19\% & -24\% & -13\%  & -20\% \\
+    CEA-Curie    & -57\%  & -35\%  & -42\% & \textbf{-53}\% & -47\%          & -53\% & -48\% & -36\%  & -46\% \\
+    Unilu-Gaia   & -62\%  & -29\%  & -31\% & -59\%          & \textbf{-60}\% & -58\% & -58\% & -33\%  & -34\% \\
     \hline
@@
 value of the period length $\Delta$ has an impact on the performance of the
-random policy. Indeed, shorter periods lead to better performance. This is perhaps
-due to the use of queuing in the scheduling process:
-When a bad policy is used, the job queue builds up. By switching more
-frequently from bad to good policies, the system never has time to build up the
-job queue. In other words, while bad decisions impact the system through an
-accumulation of waiting jobs, there is less contribution to system inertia from
-good decisions, as the number of jobs in queue can not be negative. This
-interpretation also means that randomized bandit strategies are robust by
-design, and needs more subsequent experimenting to be validated.
+random policy. Indeed, shorter periods lead to better performance.
 
@@
 Figure~\ref{fig:mosb} is the analogue of Figure~\ref{fig:mosn}, this time for
-the Bandit Feedback policy with $\Delta$ = 1 week.  Observe that here the choice of policy is much less stable. The convergence is
-extremely slow, which is expected given the variability of the reward we give
-to the bandit algorithm.
+the Bandit Feedback policy with $\Delta$ = 1 week. Observe that here the choice
+of policy is much less stable. The convergence is extremely slow, which is
+expected given the variability of the reward we give to the bandit algorithm.
 

### 2018-01-08 paper/src/paper.tex
changes to the beginning of the paper

--- 2018-01-02 paper/src/paper.tex
@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Online (Bandit) Policy Selection for EASY-Backfilling}
+\title{Online Tuning of EASY-Backfilling using Queue Reordering Policies.}
 
@@
   reordering policy for the job queue under several feedback modes. We show via
-  a comprehensive experimental validation on actual logs that noisy feedback
-  (using a simulator, which is necessarily imprecise) recovers existing
+  a comprehensive experimental validation on actual logs that periodic
+  simulation of historical data can be used to recover existing
   \textit{in-hindsight} results that allow to divide the average waiting time
-  by almost 2. Moreover, we show that bandit feedback can be used by an
-  Epsilon-greedy multi-armed bandit algorithm to cut the average waiting time
-  by 40\% without using a simulator.
+  by almost 2. This results holds even when the simulator results are noisy.
+  Moreover, we show that good performances can still be obtained without a
+  simulator, under what is called bandit feedback - when we can only observe
+  the performance of the algorithm that was picked on the live system.  Indeed,
+  a simple multi-armed bandit algorithm can reduce the average waiting
+  time by 40\%.
 
@@
 \begin{equation} \label{eq:cost-noisy} w_{\mbox{noisy}}(t;P_i) =
- \rho \sum_{j=1}^{l(\Delta_t)} \, \mbox{Wait}_j^{P_i} \end{equation}
-%
-where $\rho$ is uniformly sampled in the interval $[0.80,1.20]$, thus adding a
-+/- 20\% noise on the waiting time estimated by the simulator. The overall cost
+ \rho_t^i \sum_{j=1}^{l(\Delta_t)} \, \mbox{Wait}_j^{P_i} \end{equation}
+%
+where $\rho_t^i$ is uniformly sampled in the interval $[0.80,1.20]$, thus adding a
++/- 20\% multiplicative noise on the waiting time estimated by the simulator. The overall cost
 is then defined in the same way as before, leading to:
@@
 The experimental validation performed in this work uses a mixture of small and
-large systems from different periods. Table~\ref{tab:logs} presents the 7 logs,
-which can all be obtained from the Parallel Workload
-Archive~\cite{Feitelson20142967}. We use the ``cleaned''\cite{pwa-clean} version
-of the logs as per the archive. We impose an additional filtering step to the
-workloads\footnote{This filtering step is available in the reproducible
-workflow\cite{repro} as shell script
+large systems from different periods. These platforms are all targeted by this
+work, since they all are homogenous systems that run a batch scheduler.
+Table~\ref{tab:logs} presents the 7 logs, which can all be obtained from the
+Parallel Workload Archive~\cite{Feitelson20142967}. We use the
+``cleaned''\cite{pwa-clean} version of the logs as per the archive. We impose
+an additional filtering step to the workloads\footnote{This filtering step is
+available in the reproducible workflow\cite{repro} as shell script
 \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter|} in order to
@@
 work~\cite{jsspp17} and assume for these experiments that the system is in a
-stationary state. While we do provide guidelines as how to incorporate tracking
-variant of our algorithms in Section~\ref{sec:online} by introducing a decaying
-discount factor $\lambda$, this will not be studied in the experiments where 
-$\lambda=1$, and is left for future research.
+stationary state. We do provide guidelines as how to incorporate tracking
+into our algorithms in Section~\ref{sec:online} by introducing a decaying
+discount factor $\lambda$. Our experiments use stationary resampled data, 
+and therefore no decay is necessary: $\lambda$ will be equal to $1$.
 

### 2018-01-08 paper/src/paper.tex

--- 2018-01-07 paper/src/paper.tex
@@
 work~\cite{jsspp17} and assume for these experiments that the system is in a
-stationary state. While we do provide guidelines as how to incorporate tracking
-variant of our algorithms in Section~\ref{sec:online} by introducing a decaying
-discount factor $\lambda$, this will not be studied in the experiments where
-$\lambda=1$, and is left for future research.
+stationary state. We do provide guidelines as how to incorporate tracking
+into our algorithms in Section~\ref{sec:online} by introducing a decaying
+discount factor $\lambda$. Our experiments use stationary resampled data, 
+and therefore no decay is necessary: $\lambda$ will be equal to $1$.
 

### 2018-01-08 paper/src/paper.tex
figures change, obandit ref fix
changes

@@
 decision time and therefore, making it much more than the natural execution of
-EASY.
+EASY. The adaptation of EASY was studied under a different framework than ours
+in~\cite{talby2005improving}. Other works study the reservation
+mechanism~\cite{srinivasan2002selective}. Online portfolio approaches are also
+studied for the cloud setting~\cite{deng2013exploring}.
 
@@
     where $wait_j$ is the time job $j$ has been waiting until the time at which the decision is taken.
+    This corresponds the the value of the stretch(or slowdown) where the user runtime estimate
+    is used.
   \item SEXP: Smallest Expansion Factor First 
@@
   \item SRF: Smallest Ratio First
-  \item LAF: Largest Area $ \widetilde{p_j} \times q_j$ First
+  \item LAF: Largest Area $ \widetilde{p_j} \times q_j$ First (often called "Total Resources")
   \item SAF: Smallest Area First
@@
 where $\rho_t^i$ is uniformly sampled in the interval $[0.80,1.20]$, thus adding a
-+/- 20\% multiplicative noise on the waiting time estimated by the simulator. The overall cost
++/- 20\% multiplicative noise on the waiting time estimated by the simulator.
+This value is taken as to be sufficiently larger than values previously reported,
+see e.g. ~\cite{dutot:hal-01333471} which reports 2 percent variations of the
+simulated metric compared to the metric on the real systems.
+The overall cost
 is then defined in the same way as before, leading to:
@@
   \item For all the above strategies based on exact simulation (referred to as
-    ``Simulated feedback'' in the remainder), on noisy simulation (referred to as
-    ``Noisy Simulated Feedback'') and on $\epsilon$-greedy exploration (referred to as
-    ``Bandit Feedback''), we experiment with two policy periods: $\Delta \in
-    \left\{ 1 \text{day},1 \text{week} \right\}$. This choice is a consequence
-    of the natural rhythm of human activity: Daily or weekly simulation periods
-    seem to ease comparing of performance between two periods.
-    Indeed, arbitrary period values would needlessly increase the variability
-    of the cost metric. For the bandit feedback strategy, $\epsilon$ is set to $0.5$ and
-    $\lambda$ to $1$. This means that we do not use any decay in the estimation
-    of the overall cost. Choosing to eliminate decay in this experimental
+    ``Simulated feedback'' in the remainder), on noisy simulation (referred to
+    as ``Noisy Simulated Feedback'') and on $\epsilon$-greedy exploration
+    (referred to as ``Bandit Feedback''), we experiment with two policy
+    periods: $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$. This
+    choice is a consequence of the natural rhythm of human activity: Daily or
+    weekly simulation periods seem to ease comparing of performance between two
+    periods.  Indeed, arbitrary period values would needlessly increase the
+    variability of the cost metric. For the bandit feedback strategy, $\lambda$
+    is set to $1$. This means that we do not use any decay in the estimation of
+    the overall cost. Choosing to eliminate decay in this experimental
     validation is natural since our resampled data is in a stationary regime
-    \textit{by construction}, as explained in~\ref{ssub:resampling}. The choice
-    of $\epsilon=0.5$ is a reasonable default;
+    \textit{by construction}, as explained in~\ref{ssub:resampling}. The
+    $\epsilon$ hyperparameter will be set at a value of $0.1$ for all traces, which is the
+    result of a leave-one out majority vote selection process on
+    $[0.1,0.3,0.5,0.7,0.9]$. In this process, one trace among the 7
+    from Table~\ref{tab:logs} is taken out, and the most frequent best
+    hyperparameter value is used. Table~\ref{tab:hyperparam} outlines the
+    result of this cross-validation. Assuming workload traces are I.I.D, This
+    ensures that the reported results are fair with respect to the situation
+    where the reader of the study would implement the strategy directly.
 
@@
 \end{itemize}
+
+\begin{table}[]
+  \centering
+  \ra{1.3}
+  \caption{Hyper-parameter leave-one-out selection for $\epsilon$.}
+  \label{tab:hyperparam}
+  \begin{tabular}{@{}lr@{}}
+    \hline
+    Name          & best $\epsilon$\\
+    \hline
+    KTH-SP2       & 0.1 \\
+    CTC-SP2       & 0.1 \\
+    SDSC-SP2      & 0.1 \\
+    SDSC-BLUE     & 0.1 \\
+    ANL-Intrepid  & 0.3 \\
+    CEA-Curie     & 0.3 \\
+    Unilu-Gaia    & 0.1 \\
+    \hline
+  \end{tabular}
+\end{table}
 

### 2018-01-08 paper/paper.tex
some changes

--- 2018-01-08 paper/src/paper.tex
@@
 system and wait for the system to be empty when job submissions cease at the
-end of the submission period.
+end of the submission period. Figure~\ref{fig:resimu}
+
+\begin{figure*}[]
+  \centering
+  \input{dia/_build/algo.tex}
+
+  \caption{Illustration of the online simulation-based strategy. Every week,
+  the wokload from the past week is used to run simulations using every
+  reordering policy. The "model" $w$ is updated according
+  to~\eqref{eq:tot-cost-exact} (or~\eqref{eq:tot-cost-noisy} in the Noisy case
+  from Section~\ref{sub:noisy}) and the policy for the next week is chosen
+  according to~\eqref{eq:select-exact}}.
+
+  \label{fig:resimu}
+\end{figure*}
 
@@
 would however be too optimistic and would represent an upper bound on what can
-be achieved by a selection strategy based on simulation. In order to have a
+be achieved by a selection strategy based on simulation.
+
+In order to have a
 more realistic estimate of the behavior of a selection strategy based on
@@
 \begin{itemize}
+    \label{alg:epsilon}
 \item With probability $(1-\epsilon)$, select the policy $P$ that minimizes the cost over previous time periods (exploitation mode):
@@
 
-Several studies study the decay of $\epsilon$ over time, the exploration being
-less important once the estimates for the different policies are reliable
-\cite{Tokic:2010}. This strategy has however not been beneficial in our case.
-We believe this is due to the properties of the traces we use, as explained in
+Figure~\ref{fig:bandit} illustrates this process. Several studies study the
+decay of $\epsilon$ over time, the exploration being less important once the
+estimates for the different policies are reliable \cite{Tokic:2010}. This
+strategy has however not been beneficial in our case. We believe this is due to
+the properties of the traces we use, as explained in
 Section~\ref{sec:experiments}. The $\epsilon$-greedy strategy is known to be
@@
 simple $\epsilon$-greedy method.
+
+\begin{figure*}[]
+  \centering
+  \input{dia/_build/bandit.tex}
+
+  \caption{Illustration of the bandit-based strategy. Every week,
+  the wokload from the past week is used to run simulations using every
+  reordering policy. The "model" $w$ is updated according
+  to~\eqref{eq:tot-cost-greedy} and the policy for the next week is chosen
+  according to~\eqref{alg:epsilon}. We denote by $P_{\mbox{bandit},t}$ the 
+  policy chosen by the bandit at week $t$.
+  }.
+
+  \label{fig:bandit}
+\end{figure*}
 

### 2018-01-08 paper/run-latex.sh

@@ -1,166 +1,0 @@
-source $stdenv/setup
-
-mkdir -p $out
-
-export VARTEXFONTS=$TMPDIR/texfonts
-export TEXMFCNF=$TMPDIR:
-echo 'max_print_line = 8192' >> $TMPDIR/texmf.cnf
-
-mkdir root
-cd root
-
-startDir=$(perl $copyIncludes $includes)
-cd $startDir
-
-for i in $extraFiles; do
-  if test -d $i; then
-      ln -sf $i/* .
-  else
-      ln -sf $i $(stripHash $i)
-  fi
-done
-
-rootName=$(basename $(stripHash "$rootFile"))
-
-rootNameBase=$(echo "$rootName" | sed 's/\..*//')
-
-if test -n "$generatePDF"; then
-    latex=pdflatex
-else
-    latex=latex
-fi
-
-latexFlags="-file-line-error"
-tmpFile=$out/log
-
-showError() {
-    echo
-    echo "LATEX ERROR (LAST LOG LINES SHOWN):"
-    tail -n 20 $tmpFile
-    bzip2 $tmpFile
-    exit 1
-}
-
-runLaTeX() {
-    if ! $latex $latexFlags $rootName >$tmpFile 2>&1; then showError; fi
-    runNeeded=
-    if fgrep -q \
-        -e "LaTeX Warning: Label(s) may have changed." \
-        -e "Rerun to get citations correct." \
-        "$tmpFile"; then
-        runNeeded=1
-    fi
-}
-
-echo
-
-
-if test -n "$copySources"; then
-    cp -prd $TMPDIR/root $out/tex-srcs
-fi
-
-
-echo "PASS 1..."
-runLaTeX
-echo
-
-
-for auxFile in $(find . -name "*.aux"); do
-    # Run bibtex to process all bibliographies.  There may be several
-    # when we're using the multibib package.
-    if grep -q '\\citation' $auxFile; then
-        auxBase=$(basename $auxFile .aux)
-        if [ -e $auxBase.bbl ]; then
-            echo "SKIPPING BIBTEX ON $auxFile!"
-        else
-            echo "RUNNING BIBTEX ON $auxFile..."
-            bibtex --terse $auxBase
-            cp $auxBase.bbl $out
-            runNeeded=1
-        fi
-        echo
-    fi
-
-    # "\pgfsyspdfmark" in the aux file seems to indicate that PGF/TikZ
[…]

### 2018-01-11 paper/paper.tex
eric comments incorporated

@@
 algorithms for reducing uncertainty in problem parameters (e.g.,\ job execution
-times\cite{learningruntimes}, memory requirements, etc.). However, only few studies
+times ~\cite{learningruntimes}, memory requirements, etc.). However, only few studies
 evaluate the impact of such techniques on the performances of the resource
@@
 Both of these methods are evaluated with a specifically developed simulator,
-which is open-sourced\cite{ocst}.
+which is open-sourced ~\cite{ocst}.
 
@@
 
-Parallel job scheduling is a old studied theoretical
-problem~\cite{Frachtenberg:2009:JSS:1692356,Feitelson:2004:PJS:2128864.2128865}
+Parallel job scheduling is an old studied theoretical
+problem (~\cite{Frachtenberg:2009:JSS:1692356,Feitelson:2004:PJS:2128864.2128865})
 whose practical ramifications, varying hypotheses, and inherent uncertainty of
@@
 heuristics (and researchers to study their behavior). The two most popular
-heuristics for HPC platforms are EASY~\cite{easy} and
-Conservative~\cite{Mu'alem:2001:UPW:380314.380315} backfilling.
-
-While Conservative Backfilling offers many advantages~\cite{bfchar}, it has a
+heuristics for HPC platforms are EASY ~\cite{easy} and
+Conservative ~\cite{Mu'alem:2001:UPW:380314.380315} backfilling.
+
+While Conservative Backfilling offers many advantages ~\cite{bfchar}, it has a
 significant computational overhead and is more complex, which mainly explains
-why many of the machines of the top500 ranking~\cite{top500} rather use a
+why many of the machines of the top500 ranking ~\cite{top500} rather use a
 variant of EASY-Backfilling instead. There is a large body of work seeking to
 improve/tune EASY. Indeed, while the basic mechanism is used by some actual
-resource and job management software (most notably SLURM~\cite{SLURMdocSCHED}),
+resource and job management software (most notably SLURM ~\cite{SLURMdocSCHED}),
 this is seldom done without fine tuning by system administrators.
@@
 The original EASY mechanism refers to a First-Come-First-Serve basis.  Several
-works explore how to tune EASY by reordering waiting and/or backfilling
-queues~\cite{Tsafrir_easypp_2005}, sometimes even in a randomized
-manner~\cite{1592720}, as well as some implementations~\cite{Jackson2001}.
-However, as successful as they may be, these works do not address the
-dependency of scheduling metrics on the workload~\cite{variability}.
-These studies most often report \textit{post-hoc} performance since they
-compare algorithms after the workload is known.
-
-The dynP scheduler~\cite{streit_selftuning_2002} suggests a systematic method
-to tuning these queues. Although it requires simulated scheduling runs at
-decision time and therefore, making it much more than the natural execution of
-EASY. The adaptation of EASY was studied under a different framework than ours
-in~\cite{talby2005improving}. Other works study the reservation
-mechanism~\cite{srinivasan2002selective}. Online portfolio approaches are also
-studied for the cloud setting~\cite{deng2013exploring}.
+works explore how to tune EASY by reordering waiting and/or backfilling queues
+~\cite{Tsafrir_easypp_2005}, sometimes even in a randomized manner
+~\cite{1592720}; some implementations use this method as well
+~\cite{Jackson2001}. However, as successful as they may be, these works do not
+address the dependency of scheduling metrics on the workload
+~\cite{variability}. These studies most often report \textit{post-hoc}
+performance since they compare algorithms after the workload is known.
+
+The dynP scheduler ~\cite{streit_selftuning_2002} suggests a systematic method
+to tuning these queues. The algorithm requires simulated scheduling runs at
+decision time, therefore the time-to-decision of the scheduling is higher than
+that of EASY. The adaptation of EASY was studied under a different framework
+than ours in ~\cite{talby2005improving}. Other works study the reservation
+mechanism ~\cite{srinivasan2002selective}. Online portfolio approaches are also
+studied for the cloud setting ~\cite{deng2013exploring}.
 
@@
 works use collaborative filtering to co-locate tasks in clouds by estimating
-application interference~\cite{7516031}. Some others are closer to the
+application interference ~\cite{7516031}. Some others are closer to the
 application level. For instance,  ~\cite{fmodeling} uses binary classification
@@
 Several works use similar techniques in the context of HPC, in
-particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
-runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
+particular ~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
+runtime estimations should improve the scheduling ~\cite{chiang_impact_2002}.
 Some algorithms estimate runtime distributions model and choose jobs using
-probabilistic integration procedures~\cite{Nissimov2008}. However, these works
+probabilistic integration procedures ~\cite{Nissimov2008}. However, these works
 do not address the duality between the cumulative and maximal scheduling costs,
-as mentioned in~\cite{learningruntimes}. While these previous works intend to
+as mentioned in ~\cite{learningruntimes}. While these previous works intend to
 estimate uncertain parameters, we consider here a more pragmatic approach,
@@
 
-Existing work~\cite{jsspp17} takes this approach offlines, by
+Existing work ~\cite{jsspp17} takes this approach offline, by
 splitting workload data in a training/testing fashion. This work studies
@@
 A multi-armed bandit (MAB) problem is a sequential allocation problem with
-partially observable rewards\cite{bubnow}. At every round, an action called an
-"arm" in the litterature must be chosen from a fixed set and the corresponding
+partially observable rewards ~\cite{bubnow}. At every round an action, called an
+"arm" in the literature, must be chosen from a fixed set and the corresponding
 reward is observed. The goal of a MAB algorithm is to maximize the total reward
@@
 are the (original) stochastic setting and the adversarial setting of regret
-minimization. See~\cite{thompson} for the original work on the stochastic
-setting and~\cite{Auer2002} for the ''Upper Confidence Bound`` family of
-algorithms. To the best of our knowledge, the work of~\cite{Banos} is the
-earliest for the adversarial setting; see~\cite{nonstoch} for the
-``Exponential-weight'' family of algorithms. We refer to~\cite{bubnow} for a
+minimization. See ~\cite{thompson} for the original work on the stochastic
+setting and ~\cite{Auer2002} for the ''Upper Confidence Bound`` family of
+algorithms. To the best of our knowledge, the work of ~\cite{Banos} is the
+earliest for the adversarial setting; see ~\cite{nonstoch} for the
+``Exponential-weight'' family of algorithms. We refer to ~\cite{bubnow} for a
 comprehensive review of the field. While these algorithms bound the cumulative
@@
 The simple heuristic called \textit{Epsilon-Greedy} introduced
-in~\cite{Auer2002} does not have this requirement. For this reason, this
+in ~\cite{Auer2002} does not have this requirement. For this reason, this
 heuristic will be used in this paper and is described in
@@
 concurrent parallel jobs with rigid (known and fixed) resource requirements on
-a HPC platform represented by a pool of $m$ identical resources. This is an
+an HPC platform represented by a pool of $m$ identical resources. This is an
 on-line problem since the jobs are submitted over time and their
@@
    \item Requested running time $\widetilde{p_j}$ (sometimes called
-     \textit{walltime}), which is an upper bound of $p_j$.
+     \textit{walltime}), which is in general an upper bound of $p_j$.
 
@@
 the most widely used policy due to its simple and robust implementation and
-known benefits such as high system utilization~\cite{easy}. This strategy has
+known benefits such as high system utilization ~\cite{easy}. This strategy has
 no worst case guarantee beyond the absence of starvation (i.e.\ every job will
@@
 the more commonly used objectives. The waiting time of a job is 
-
  \begin{equation}
@@
  \end{equation}
-
 where $start_j$ is the starting time of a job.
@@
 waiting time. It is worth noting that other metrics such as the maximum waiting
-time of all the jobs are also worthy of interest. Unfortunately, this criteria
+time of all the jobs are also worthy of interest. Unfortunately, this criterion
 is antagonistic in nature with the average waiting time. Section~\ref{sub:th}
@@
 queue is ordered in FCFS order in the original heuristic, it is possible to
-reorder it at will. We focus on a search space of 10 reordering policies.
+reorder it at will. We focus on a search space of 12 reordering policies.
 
 \begin{enumerate}
-  \item FCFS: First-Come First-Serve, which is the most commonly used policy~\cite{easy}.
+  \item FCFS: First-Come First-Serve, which is the most commonly used policy ~\cite{easy}.
   \item LCFS: Last-Come First-Serve.
@@
   \item SQF: Smallest resource requirement First.
-  \item LEXP: Largest Expansion Factor First~\cite{bfchar}, where the expansion
+  \item LEXP: Largest Expansion Factor First ~\cite{bfchar}, where the expansion
     factor is defined as follows:
@@
 possible without making any judgement on which policy should perform well in
-practice. It does include most policies from related works in this area that we
+practice. It does include most policies from related works that we
 are aware of. In the following, we denote these policies by $P_i$ with $i = 1
@@
 
-Reordering the job queue means losing the no-starvation guarantee and some
+Reordering the job queue means losing the no-starvation guarantee; some
 individual jobs therefore can wait an undue amount of time. It is possible to
@@
 between consecutive periods. This assumption is validated in the study
-presented in~\cite{jsspp17}. It further entails that the behavior of a policy
+presented in ~\cite{jsspp17}. It further entails that the behavior of a policy
 during the previous periods reflects its behavior on the current one, so that
@@
 be some variability between different periods for certain cost
-metrics~\cite{feitelson2001metrics}. This is illustrated in
+metrics ~\cite{feitelson2001metrics}. This is illustrated in
 Figure~\ref{fig:variability} which displays the average waiting time of various
@@
 space to consider is infinite. This is not prohibitive \textit{per se}, as
-modern methods~\cite{aprl} can bypass dimensionality issues via function
-approximation. However, these methods rely on large amounts of data, and we are
-studying online methods\footnote{The reinforcement learning literature uses the
-term ``\textit{on-policy}'' for these methods.}. We rely here on simpler, yet we
-believe more effective, strategies to solve this problem.  These strategies are
-applied online and rely on exact simulation, noisy simulation and
-$\epsilon$-greedy bandit exploration.
+modern methods ~\cite{aprl} can bypass dimensionality issues via function
+approximation. However, these methods rely on large amounts of batch data, and
+we are facing an online problem with no pre-existing data\footnote{The
+reinforcement learning literature uses the term ``\textit{on-policy}'' for this
+setting.}. We rely here on simpler, yet we believe more effective, strategies
+to solve this problem.  These strategies are applied online and rely on exact
+simulation, noisy simulation and $\epsilon$-greedy bandit exploration.
 
@@
 %
-\begin{equation} \label{eq:cost-exact} w_{\mbox{exact}}(t;P_i) =
-\sum_{j=1}^{l(\Delta_t)} \mbox{Wait}_j^{P_i} \end{equation}
+\begin{equation} \label{eq:cost-exact} w(t;P_i) = \sum_{j=1}^{l(\Delta_t)}
+\mbox{Wait}_j^{P_i} \end{equation}
 %
@@
 \begin{equation} \label{eq:tot-cost-exact} w(\rightarrow T;P_i) =
-\sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{exact}}(t;P_i) \end{equation}
+\sum_{t=0}^{T-1} \lambda^{T-1-t} w(t;P_i) \end{equation}
 %
 where $\lambda \in [0,1]$ is a decaying discount factor that can be used to
-privilege recent history (\textit{i.e.} recent periods). One then selects the
+privilege recent history (\textit{i.e.} recent periods). The above formulation
+is general and covers several possible situations : From stationary job
+distributions where $\lambda$ can be set to 1 to non-stationary distributions
+where the more recent history has to be privileged. As explained in Section
+6.1, in our experiments $\lambda$ is set to 1 as the distributions considered
+are stationary but the reader has to bear in mind that our framework can be use
+on different distributions by resorting to different values of lambda that can
+be set by cross-validation.
+
+One then selects the
 policy $P$ that minimizes the above cost for the period $\Delta_T$:
@@
 system and wait for the system to be empty when job submissions cease at the
-end of the submission period. Figure~\ref{fig:resimu}
+end of the period (see Figure~\ref{fig:resimu}).
 
@@
 
-In order to have a
-more realistic estimate of the behavior of a selection strategy based on
-simulation, we introduce noise in the simulator, as described below.
+In order to have a more realistic estimate of the behavior of a selection
+strategy based on simulation, we introduce noise in the simulator, as described
+below.
 
@@
 
-In order to simulate how the simulation strategy for selecting policies would
-work on non simulated traces, we randomly introduce noise in the estimate of
-the policy cost defined by Eq.~\eqref{eq:cost-exact} by rescaling it, either
-down or up:
-%
-\begin{equation} \label{eq:cost-noisy} w_{\mbox{noisy}}(t;P_i) =
- \rho_t^i \sum_{j=1}^{l(\Delta_t)} \, \mbox{Wait}_j^{P_i} \end{equation}
-%
-where $\rho_t^i$ is uniformly sampled in the interval $[0.80,1.20]$, thus adding a
-+/- 20\% multiplicative noise on the waiting time estimated by the simulator.
-This value is taken as to be sufficiently larger than values previously reported,
-see e.g. ~\cite{dutot:hal-01333471} which reports 2 percent variations of the
-simulated metric compared to the metric on the real systems.
-The overall cost
-is then defined in the same way as before, leading to:
-%
-\begin{equation} \label{eq:tot-cost-noisy} w(\rightarrow T;P_i) =
+We take another step to evaluate how the simulation strategy for selecting
+policies would work in a batch scheduler. Simulation of real systems can be
+imprecise. While existing work reports values as low as 2\% imprecision we take
+a conservative approach and consider imprecision up to of 20\%. We randomly
+introduce multiplicative noise in the estimate of the policy cost defined by
+Eq.~\eqref{eq:cost-exact} by rescaling it, either down or up:
+%
+\begin{equation} \label{eq:cost-noisy} w_{\mbox{noisy}}(t;P_i) = \rho_t^i
+\sum_{j=1}^{l(\Delta_t)} \, \mbox{Wait}_j^{P_i} \end{equation}
+%
+where $\rho_t^i$ is uniformly sampled in the interval $[0.80,1.20]$, thus
+adding a +/- 20\% multiplicative noise on the waiting time estimated by the
+simulator.  This value is taken as to be sufficiently larger than values
+previously reported, see e.g. ~\cite{dutot:hal-01333471} which reports 2
+percent variations of the simulated metric compared to the metric on the real
+systems.  The overall cost is then defined in the same way as before, leading
+to:
+%
+\begin{equation} \label{eq:tot-cost-noisy} w_{\mbox{noisy}}(\rightarrow T;P_i) =
 \sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i) \end{equation}
 %
-As before (see Eq.~\ref{eq:select-exact}), the policy that minimizes the above
+As before the policy that minimizes the above
 cost is selected for the period $\Delta_T$.
+
+\begin{equation} \label{eq:select-noisy} P = \mbox{argmin}_{P_i, 1 \le i \le
+12} w_{\mbox{noisy}}(\rightarrow T;P_i) \end{equation}
 
@@
 $\epsilon$-greedy exploration method, standard in reinforcement learning
-\cite{Watkins:1989} and bandit problems~\cite{Auer2002}, to trade-off
+~\cite{Watkins:1989} and bandit problems ~\cite{Auer2002}, to trade-off
 exploitation and exploration, and relies, as before, on past estimates of the
@@
 %
-\begin{equation} \label{eq:tot-cost-greedy} w(\rightarrow T;P_i) =
+\begin{equation} \label{eq:tot-cost-greedy} w_{\mbox{epsilon}}(\rightarrow T;P_i) =
   \frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta^{'}_t)} \sum_{t=0}^{T-1}
@@
 \label{eq:select-greedy}
-P = \mbox{argmin}_{P_i, 1 \le i \le 12} w(\rightarrow T;P_i)
+P = \mbox{argmin}_{P_i, 1 \le i \le 12} w_{\mbox{epsilon}}(\rightarrow T;P_i)
 \end{equation}
@@
 decay of $\epsilon$ over time, the exploration being less important once the
-estimates for the different policies are reliable \cite{Tokic:2010}. This
+estimates for the different policies are reliable ~\cite{Tokic:2010}. This
 strategy has however not been beneficial in our case. We believe this is due to
@@
 outperformed in terms of regret by other classical algorithms such as the UCB
-or EXP family of algorithms. See~\cite{bubnow} for a precise definition and
+or EXP family of algorithms (see ~\cite{bubnow}). See ~\cite{bubnow} for a precise definition and
 thorough treatment of this subject. These algorithms however all rely on
@@
 Table~\ref{tab:logs} presents the 7 logs, which can all be obtained from the
-Parallel Workload Archive~\cite{Feitelson20142967}. We use the
-``cleaned''\cite{pwa-clean} version of the logs as per the archive. We impose
+Parallel Workload Archive ~\cite{Feitelson20142967}. We use the
+``cleaned'' ~\cite{pwa-clean} version of the logs as per the archive. We impose
 an additional filtering step to the workloads\footnote{This filtering step is
-available in the reproducible workflow\cite{repro} as shell script
+available in the reproducible workflow ~\cite{repro} as shell script
 \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter|} in order to
@@
 This is the approach used by high-fidelity batch scheduling simulators such as
-BatSim~\cite{batsim}. In the present work, the focus is on studying the
+BatSim ~\cite{batsim}. In the present work, the focus is on studying the
 EASY-Backfilling mechanism in itself, without addressing the allocation
@@
 any available processor on the system. During this work, we developed a
-lightweight simulator\cite{ocst} and now release it to the community along with
+lightweight simulator ~\cite{ocst} and now release it to the community along with
 this article. See the reproducibility paragraph below for more information.
@@
 Comparing algorithms in batch scheduling is made especially difficult by the
-fluctuations in commonly used metrics~\cite{feitelson1998metrics}. The variability in
+fluctuations in commonly used metrics ~\cite{feitelson1998metrics}. The variability in
 performance is known to outrank the available sample sizes: \textit{It is not
@@
 conclusions about which algorithm is better. This is a known
-problem~\cite{flurries} that motivated decades of research in workload
-modeling~\cite{feitbook}. Workload models have drawbacks in that they usually
+problem ~\cite{flurries} that motivated decades of research in workload
+modeling ~\cite{feitbook}. Workload models have drawbacks in that they usually
 lose the details of the workloads. This is problematic due to the non-linear
 aspect of the job scheduling process. This is discussed
-in~\cite{feitresampling}, an approach which attempts to solve this problem by
+in ~\cite{feitresampling}, an approach that attempts to solve this problem by
 resampling data directly from the original traces. In this work, we use a
@@
 us to preserve the dependency within a week. Although the system state may not
-be independent from one week to an other, we rely on existing
-work~\cite{jsspp17} and assume for these experiments that the system is in a
+be independent from one week to an other, as illustrated in~\cite{jsspp17},
+it is reasonable
+to assume that the system is in a stationary state.~\cite{jsspp17} and assume for these experiments that the system is in a
 stationary state. We do provide guidelines as how to incorporate tracking
 into our algorithms in Section~\ref{sec:online} by introducing a decaying
-discount factor $\lambda$. Our experiments use stationary resampled data, 
-and therefore no decay is necessary: $\lambda$ will be equal to $1$.
+discount factor $\lambda$. In our experiments, we use stationary resampled data, 
+and therefore no decay is necessary: $\lambda$ here is set to $1$.
 
@@
     hyperparameter value is used. Table~\ref{tab:hyperparam} outlines the
-    result of this cross-validation. Assuming workload traces are I.I.D, This
+    result of this cross-validation. Assuming workload traces are i.i.d, this
     ensures that the reported results are fair with respect to the situation
@@
 
-We publish a declarative workflow~\cite{repro} based on our
-simulator~\cite{ocst} that runs experiments and generates all the figures from
+We publish a declarative workflow ~\cite{repro} based on our
+simulator ~\cite{ocst} that runs experiments and generates all the figures from
 this paper.
@@
 packaging software. As our experiments are CPU bound, we decide to opt for the
-software packaging approach~\cite{nix}. This allows to replicate the experiments on
+software packaging approach ~\cite{nix}. This allows to replicate the experiments on
 clusters where virtualization is not available or kernels are too old for
@@
 
-All the code associated with this work is released under the ISC~\cite{isc}
+All the code associated with this work is released under the ISC ~\cite{isc}
 license.
@@
 All data from this paper were obtained from the Swf Parallel Workloads
-Archive~\cite{Feitelson20142967}.
-
- All experiments are tied together using zymake~\cite{zymake}, a minimalistic
+Archive ~\cite{Feitelson20142967}.
+
+ All experiments are tied together using zymake ~\cite{zymake}, a minimalistic
  workflow system designed for computational experiments\footnote{The zymake system
@@
  \begin{itemize}
-   \item Data. Data extraction from archives, filtering, resampling.
+   \item Data extraction from archives, data filtering, and data resampling.
      This is principally using shell scripts for data manipulation (see file
@@
    \item Simulation. This step runs the lightweight ocaml backfilling
-     simulator specially written for this work. this simulator is made
-     available under the ISC~\cite{isc} license as a persistent zenodo
-     archive at~\cite{ocst}.
-
-   \item Analysis. This step runs R code that generates the figures
-     presented in this paper.
+     simulator specially written for this work. This simulator is made
+     available under the ISC ~\cite{isc} license as a persistent zenodo
+     archive at ~\cite{ocst}.
+
+   \item Analysis of the results. This step runs R code that generates the
+     figures presented in this paper.
  \end{itemize}
@@
 
-So that the values appearing in Table~\ref{tab:full} are exactly:
+The values appearing in Table~\ref{tab:full} are exactly:
 

### 2018-01-11 paper/revision.tex
revision letter

\documentclass[]{article}

\usepackage{listings}
\usepackage{fullpage}
\usepackage{dirtree}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bbm}
\usepackage[scientific-notation=true]{siunitx}

\usepackage{amssymb}
\usepackage{tabulary}

\usepackage{etoolbox}

\usepackage{algorithm}
\usepackage{algorithmic}

\title{Major revision comments.}

\begin{document}

\maketitle

\section{Reviewer 1}

\textbf{The manuscript describes an experimental study to modify a scheduling
approach common in many large parallel computer systems. The authors propose to
replace the well-established First-Come-First-Serve approach combined  with
EASY-backfilling by a strategy using reordering of the queue. To prevent job
starvation, the authors resort to FCFS for a job as soon as the waiting time of
this job exceeds a threshold. Although this threshold approach is well known,
the authors do not provide any references of previous publications using this
approach. }

\textbf{The reviewer sees significant shortcomings in the experimental design
of the manuscript. In general, computation experiments should follow the
guidelines of experiments in natural sciences, that is, an experiment has the
goal to prove or disprove a given hypothesis. The authors do not describe such
hypothesis nor can the reviewer extract it from the manuscript (except for a
trivial hypothesis: we can beat EASY-Backfilling for some traces). }

\textbf{Instead, the authors present a simple trial-and error study that
exploits some parts of the problem space without justifying the restriction of
the problem space. For instance, the authors suggest 12 reordering policies
(Section 4.1) with the justification of the selecting a search space "as
semantically diverse as possible" instead of carefully arguing which reordering
policy has the potential to improve performance. For the scheduling objective,
the authors use "the waiting times of jobs, which is one of the more commonly
used objectives" although they state before that a "system administrator may
use one or multiple cost metric(s)". Although this approach is the opposite of
the approach used for the reordering policies, the authors do not justify the
design decisions of their study. }

\textbf{Similarly, the authors simply use 40h as threshold and simply state
that the "choice of this default stems from the analysis developed in" a former
manuscript of the authors that is so far unpublished. Such approach is simply
unacceptable in view of the reviewer.}

This is indeed a mistake on our part which made the way to the final version,
this is now corrected. The article is "Tuning Backfilling Queues" from JSSPP17.\\

\textbf{The authors use 7 traces recorded over a time span of 18 years. They do
not discuss whether these traces a comparable with respect to their study
despite large differences in the number of processors, jobs and trace lengths.}

\textbf{In their algorithmic description, the authors include a decaying factor
$\lambda$ although they later simply state "this will not be studied in the
experiments where $\lambda=1$ and is left for future research". Therefore, the
authors indicate that even in their own opinion, research regarding this
algorithm is not complete but the presented findings are only a first step. The
reviewer feels that such first step may be helpful in a workshop to discuss
directions to complete such study but not for reporting the results of a study
in a journal. }

\textbf{Altogether, the reviewer believes that an experimental study with so
many design problems is not relevant to the readers of TDPS. }

\textbf{The reviewer does not see how the authors can overcome these
shortcomings even in a major revision and therefore suggest rejection of the
manuscript.}

\textbf{In addition, the reviewer sees several minor problems in the manuscript
that the authors can correct in a major revision. The reviewer only briefly
describes such problems in the following:}

\textbf{Abstract: Any reader of TPDS can expect to understand the abstract
without having to use additional literature. Therefore, the abstract should
only include expressions that are common knowledge: it is at least doubtful
that every reader understands the used abbreviations and expressions like
"Epsilon-greedy multi armed bandit algorithm". }

\textbf{Previous publications explain that studies using traces suffer from
fill-ins in the starting period and from draining in the final period. The
authors mention this problem but do not discuss their impact and strategies to
avoid these problems although other publications present such strategies.
Previous studies have described that the KTH trace is hardly comparable with
other traces due to a specific acceptance policy. The authors do not discuss
this problem. The authors use either 1 week or 1 day feedback justifying their
choices with the periodicity of the traces. Although such approach is
reasonable, it is a hypothesis and requires validation by using other feedback
periods as well and comparing the results. }

\textbf{There are several publications discussing the disadvantages in using
traces and suggesting workload models instead. The authors do not consider
these studies in their manuscript without explaining why.}

We leave this discussion to other works, in particular, we suggest the EuroPar
article 'Workload Resampling for Performance Evaluation of Parallel Job
Schedulers'.\\

\textbf{Problems of the introduction explaining the goal of the study are
likely related to the already mentioned design problems of the study. The same
holds for the organization of the manuscript.}

\textbf{The authors can improve readability of the manuscript by explaining the
underlying algorithmic approach in some more detail.}

We added two figures to illustrate the algorithms.\\

\textbf{Since the journal provides a platform for supplemental material, the
reviewer does not understand why readers must contact the authors by email to
obtain data to reproduce the study.  }

As previously concerning the missing reference, this was inadvertent on our
part! This is fixed in the major revision. Apologies for this error.

\section{Reviewer 2}

\textbf{I basically liked this paper and support its publication.  The
following comments make some suggestions for possible improvements.}

\textbf{Title: I would suggest to remove the "(bandit)" from the title, and
instead use something like "Online Queue-Ordering Policy Selection for
EASY-Backfilling" -- emphasize what you are doing, not how.}

The need for more careful framing of the goals of the study was stressed by all
reviewers. Accordingly, we made the scope of the study more clear and changed
the title, to "Online Tuning of EASY-Backfilling using Queue Reordering
Policies."\\

\textbf{Related work: additional papers you may want to look at are}
\begin{enumerate}

  \item \textbf{"Exploring portfolio scheduling for long-term execution of
    scientific workloads in IaaS clouds", from SC'13, suggests how to select
    from a large portfolio of possible schedulers;}

  \item \textbf{"Improving and stabilizing parallel computer performance using
    adaptive backfilling", from IPDPS 2005, which also suggests online
    selection from a set of 4 alternative schedulers;}

  \item \textbf{"Selective reservation Strategies for Backfill Job Scheduling",
    from JSSPP 2002, which suggests online decisions of whether to make
    reservations for skipped jobs - similar effect to your thresholding
    scheme;}

  \item \textbf{"Self-tuning systems", from IEEE Software 1999, which is about
    simulating performance with local workloads to select systems parameter
    values.}

\end{enumerate}
Thank you for your suggestions, we had not come across these papers! There are indeed very related to our effort.

\textbf{Section 2.2 - could be good to give a short description of how a basic
    version of such an algorithm works, instead of just noting a list of
    versions.}

We did add a basic description of the setting in the description.\\

\textbf{Section 3.1 notation - intuitively p could represent processors just as
well as processing time, so I would suggest $t_i$ for time and $p_i$ for
processors.  Similarly, in Section 4.1 I for one would find the names more
intuitive with SJF/LJF instead of SPF/LPF (shortest/longes job first) SPF/LPF
instead of SQF/LQF (smallest/largest processors first) expansion factor is
often called slowdown area is often called total resources}

We agree that $p_i$ is not very intuitive, but for lack of a standard we chose
to use notation from scheduling theory.  In face of the many possible acronyms,
we chose to rather be consistent with the $p_i$ and $q_i$ notation - SPF ist
smallest $p_i$ first and SQF is smallest $q_i$ first..  We added the remark
concerning the total resources and the expansion factor. This notation comes
from the "Characterization of backfilling strategies for parallel job
scheduling" article, where the authors call Expansion Factor the value of the
slowdown where the upper bound on the running-time is used.\\

\textbf{Section 5.1 exact simulation: I understand why you divide the time into
    periods as part of online tuning. But I disagree with doing so also in the
    simulations - as you yourselves say, it just causes problems with boundary
    effects.  And you can discount more distant jobs without using periods.}

We agree that this design choice can be discussed. Our decision to divide
makes the study more realistic with respect to the cases where simulations are
expensive(which is most large systems).  Indeed, the more realistic simulations
are, the more time they take. For example, precise simulation frameworks such
as SimGrid/Batsim have high runtimes, and simulating the history of the whole
system every period is prohibitive if the network topology and/or multiple
resource constraints are used. As for discounting - it can indeed be used
in both cases.\\

\textbf{Section 5.2 - adding noise to the output rather than to the input
    workload doesn't make sense in my opinion. And the result that it gave the
    same results is totally expected - you add random noise with mean 0 and
    then fined that the average didn't change.}

We do agree that this does not change the estimate - the point of adding 
this noise was to show that the sample size is not prohibitive in practice.
Indeed, we learn the policy almost as fast with this noise, which
motivates the use of the technique in real-world cases. The point of this noise
is really to compare the difference in the shapes of the two green curves
in Figure 3.\\

\textbf{Section 6.1.1 traces - I think negative values are just -1 meaning no data is available.}

You are right - and we made this point clear in the text.\\

\textbf{Section 6.3.1 - I think you can say more about the post-factum results.  Even
if it is not your focus, it is still interesting that for example ordering
by size is more beneficial than by time, and that it should be
largest-first (and the SJF-like scheme is not good).}

Doing this would drown the point in the author's opinion. We agree it is an
interesting point - we now refer to our previous publication 'Tuning
Backfilling Queues'(JSSPP17), which was inadvertently ommited from the related
works of the first submission.\\

  \textbf{Section 6.3.2 FCFS vs. Random: you justify the result be a conjecture that jobs do not pile up in the queue.  you should verify this by counting the jobs int he queue in the simulation.}

We aggree that this conjecture is not warranted - we removed it in the revision.
We do have plans to explore this aspect for a future publication where the
discussion will be more appropriate.\\

\textbf{Table 3: add columns with best static result for comparison.}

Done, thanks you for the suggestion.\\

\textbf{Description of Figure 6 results: this is not slow convergence, it is no
convergence.  and using epsilon=0.5 is much too high: it means that half of the
time you are choosing randomly, and that's why all of the possible approaches
are used approximately the same fraction of the time. What you want is to
initially use them all, but then reduce epsilon with time till you reach some
reasonable value like random exploration in 1 of 10 or 1 of 20 periods.}

We changed the experimental protocol to use leave-one-out cross validation,
which did improve the bandit results. The selected exploration rate was 0.1 for
all traces, although it is not the best choice for all of them. (see the new
Table 2) The choice of a fixed exploration rate is one of simplicity - while
decaying exploration does provide better performance, it is also harder to
tune.\\

\textbf{Reference [4] is often referred to and gives important background, but
is not available. NOTE: this journal does not really do blind review. Your
names are given on the first page.}

This is corrected in the major revision - we inadvertently ommited this
when preparing the submission. The reference is 'Tuning Backfilling Queues'(JSSPP17).\\

\textbf{Typos and usage [...]}

We incorporated these changes and thank the reviewer for their detailed read.\\

\section{Reviewer 3}

\textbf{I think that the current experiments are not sufficient to show the advantage of the proposed strategy.
The authors compare the performance of the proposed strategy with the baseline performance derived by EASY Backfilling. The performance of EASY Backfilling is good metric as the baseline, but comparison with EASY Backfilling is not enough to discuss the advantage of the proposed strategy. A lot of work about tuned Backfilling has been presented as discussed in Sec. 2. The author should discuss the performance improvement compared with the existing tuned backfilling strategies.
The experimental results in Table 2 show that LAF and LQF show best performance among 12 reordering policies. This indicates that we can expect significant performance improvement by using static strategies, Backfilling with LAF or that with LQF, compared with EASY Backfilling. If the performance gap between the proposed strategy and the static strategies are small, the static strategies may be preferable because they are less complicated and run with smaller scheduling overhead. A comparison between the proposed strategy and the static strategies is needed to show the advantage of the proposed strategy.}

We do agree with the reviewer that in order to really motivate the use of
bandit/resimulation aproaches, one should compare the performance of static
policies, say using leave-one-trace-out cross validation, to the performance of
the adaptive variants, using leave-one-trace-out cross validation for their
hyperparameters.
However, this work does not aim to underline the necessity of the bandit
approach (the title has been changed and the introduction modified in the major
revision to reflect this better). Rather, we try to show its advantages,
feasibility and limitations. 
There are two main reasons for not trying to be too demonstrative of the superiority
of the adaptive approach using our hypotheses.

First, this study uses a limited amount of workload logs. In consequence, the
process of comparing the adaptive (hyperparameter-cross-validated) choice to
the static (cross-validated in the set of fixed policies) choice would not be
really conclusive.  Indeed, the more workload logs are presented, the more
diversity is present and the more adaptation is necessary.
Second, we use a simplistic simulation model. In this model, we do make
machines *more* similar to each other by neglecting topological differences. It
can be reasonably expected that differences between systems should *increase*
rather than decrease, when moving away from a simplistic model to the real
systems.
We modified the present experiments to use leave-one-out cross
validation in the hyperparameter selection, and made the reasons for this
choice of experimental protocol clearer in the text.
We are preparing a subsequent study using an approach closer to what you
describe, in which we greatly increase the number of workload logs used.\\

\textbf{The authors assume +/-20% noise in the simulation running in the
proposed strategy. However, the reason why this assumption is suitable is not
presented. The authors should discuss the noise of 20% is suitable to represent
real world jobs.}

A justification was indeed missing, this value is taken as to be sufficiently
larger than values previously reported, see e.g. Batsim: a Realistic
Language-Independent Resources and Jobs Management Systems Simulator which
reports 2 percent variations of the simulated metric compared to the metric on
the real systems. We added this remark to the text.\\

\textbf{Minor comments for improving presentation}

We incorporated these changes and thank the reviewer for their detailed read.\\

\bibliographystyle{abbrv}
\bibliography{bibliography}

\end{document}

### 2018-01-12 paper/paper.tex
paper fix

@@
 algorithms for reducing uncertainty in problem parameters (e.g.,\ job execution
-times ~\cite{learningruntimes}, memory requirements, etc.). However, only few studies
+times~\cite{learningruntimes}, memory requirements, etc.). However, only few studies
 evaluate the impact of such techniques on the performances of the resource
@@
 Both of these methods are evaluated with a specifically developed simulator,
-which is open-sourced ~\cite{ocst}.
+which is open-sourced~\cite{ocst}.
 
@@
 heuristics (and researchers to study their behavior). The two most popular
-heuristics for HPC platforms are EASY ~\cite{easy} and
-Conservative ~\cite{Mu'alem:2001:UPW:380314.380315} backfilling.
-
-While Conservative Backfilling offers many advantages ~\cite{bfchar}, it has a
+heuristics for HPC platforms are EASY~\cite{easy} and
+Conservative~\cite{mualem_utilization_2001} backfilling.
+
+While Conservative Backfilling offers many advantages~\cite{bfchar}, it has a
 significant computational overhead and is more complex, which mainly explains
-why many of the machines of the top500 ranking ~\cite{top500} rather use a
+why many of the machines of the top500 ranking~\cite{top500} rather use a
 variant of EASY-Backfilling instead. There is a large body of work seeking to
 improve/tune EASY. Indeed, while the basic mechanism is used by some actual
-resource and job management software (most notably SLURM ~\cite{SLURMdocSCHED}),
+resource and job management software (most notably SLURM~\cite{SLURMdocSCHED}),
 this is seldom done without fine tuning by system administrators.
@@
 
-The dynP scheduler ~\cite{streit_selftuning_2002} suggests a systematic method
+The dynP scheduler~\cite{streit_selftuning_2002} suggests a systematic method
 to tuning these queues. The algorithm requires simulated scheduling runs at
@@
 that of EASY. The adaptation of EASY was studied under a different framework
-than ours in ~\cite{talby2005improving}. Other works study the reservation
-mechanism ~\cite{srinivasan2002selective}. Online portfolio approaches are also
-studied for the cloud setting ~\cite{deng2013exploring}.
+than ours in~\cite{talby2005improving}. Other works study the reservation
+mechanism~\cite{srinivasan2002selective}. Online portfolio approaches are also
+studied for the cloud setting~\cite{deng2013exploring}.
 
@@
 works use collaborative filtering to co-locate tasks in clouds by estimating
-application interference ~\cite{7516031}. Some others are closer to the
+application interference~\cite{7516031}. Some others are closer to the
 application level. For instance, ~\cite{fmodeling} uses binary classification
@@
 Several works use similar techniques in the context of HPC, in
-particular ~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
-runtime estimations should improve the scheduling ~\cite{chiang_impact_2002}.
+particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
+runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
 Some algorithms estimate runtime distributions model and choose jobs using
-probabilistic integration procedures ~\cite{Nissimov2008}. However, these works
+probabilistic integration procedures~\cite{Nissimov2008}. However, these works
 do not address the duality between the cumulative and maximal scheduling costs,
-as mentioned in ~\cite{learningruntimes}. While these previous works intend to
+as mentioned in~\cite{learningruntimes}. While these previous works intend to
 estimate uncertain parameters, we consider here a more pragmatic approach,
@@
 
-Existing work ~\cite{jsspp17} takes this approach offline, by
+Existing work~\cite{jsspp17} takes this approach offline, by
 splitting workload data in a training/testing fashion. This work studies
@@
 A multi-armed bandit (MAB) problem is a sequential allocation problem with
-partially observable rewards ~\cite{bubnow}. At every round an action, called an
+partially observable rewards~\cite{bubnow}. At every round an action, called an
 "arm" in the literature, must be chosen from a fixed set and the corresponding
@@
 are the (original) stochastic setting and the adversarial setting of regret
-minimization. See ~\cite{thompson} for the original work on the stochastic
-setting and ~\cite{Auer2002} for the ''Upper Confidence Bound`` family of
-algorithms. To the best of our knowledge, the work of ~\cite{Banos} is the
-earliest for the adversarial setting; see ~\cite{nonstoch} for the
-``Exponential-weight'' family of algorithms. We refer to ~\cite{bubnow} for a
+minimization. See~\cite{thompson} for the original work on the stochastic
+setting and~\cite{Auer2002} for the ''Upper Confidence Bound`` family of
+algorithms. To the best of our knowledge, the work of~\cite{Banos} is the
+earliest for the adversarial setting; see~\cite{nonstoch} for the
+``Exponential-weight'' family of algorithms. We refer to~\cite{bubnow} for a
 comprehensive review of the field. While these algorithms bound the cumulative
@@
 The simple heuristic called \textit{Epsilon-Greedy} introduced
-in ~\cite{Auer2002} does not have this requirement. For this reason, this
+in~\cite{Auer2002} does not have this requirement. For this reason, this
 heuristic will be used in this paper and is described in
@@
 the most widely used policy due to its simple and robust implementation and
-known benefits such as high system utilization ~\cite{easy}. This strategy has
+known benefits such as high system utilization~\cite{easy}. This strategy has
 no worst case guarantee beyond the absence of starvation (i.e.\ every job will
@@
 \begin{enumerate}
-  \item FCFS: First-Come First-Serve, which is the most commonly used policy ~\cite{easy}.
+  \item FCFS: First-Come First-Serve, which is the most commonly used policy~\cite{easy}.
   \item LCFS: Last-Come First-Serve.
-  \item SPF: Smallest estimated Processing time $\widetilde{p_{j}}$ First ~\cite{bfchar}.
+  \item SPF: Smallest estimated Processing time $\widetilde{p_{j}}$ First~\cite{bfchar}.
   \item LPF: Longest estimated Processing time First.
@@
   \item SQF: Smallest resource requirement First.
-  \item LEXP: Largest Expansion Factor First ~\cite{bfchar}, where the expansion
+  \item LEXP: Largest Expansion Factor First~\cite{bfchar}, where the expansion
     factor is defined as follows:
@@
 between consecutive periods. This assumption is validated in the study
-presented in ~\cite{jsspp17}. It further entails that the behavior of a policy
+presented in~\cite{jsspp17}. It further entails that the behavior of a policy
 during the previous periods reflects its behavior on the current one, so that
@@
 be some variability between different periods for certain cost
-metrics ~\cite{feitelson2001metrics}. This is illustrated in
+metrics~\cite{feitelson2001metrics}. This is illustrated in
 Figure~\ref{fig:variability} which displays the average waiting time of various
@@
 space to consider is infinite. This is not prohibitive \textit{per se}, as
-modern methods ~\cite{aprl} can bypass dimensionality issues via function
+modern methods~\cite{aprl} can bypass dimensionality issues via function
 approximation. However, these methods rely on large amounts of batch data, and
@@
 simulator.  This value is taken as to be sufficiently larger than values
-previously reported, see e.g. ~\cite{dutot:hal-01333471} which reports 2
+previously reported, see e.g.~\cite{dutot:hal-01333471} which reports 2
 percent variations of the simulated metric compared to the metric on the real
@@
 $\epsilon$-greedy exploration method, standard in reinforcement learning
-~\cite{Watkins:1989} and bandit problems ~\cite{Auer2002}, to trade-off
+~\cite{Watkins:1989} and bandit problems~\cite{Auer2002}, to trade-off
 exploitation and exploration, and relies, as before, on past estimates of the
@@
 decay of $\epsilon$ over time, the exploration being less important once the
-estimates for the different policies are reliable ~\cite{Tokic:2010}. This
+estimates for the different policies are reliable~\cite{Tokic:2010}. This
 strategy has however not been beneficial in our case. We believe this is due to
@@
 outperformed in terms of regret by other classical algorithms such as the UCB
-or EXP family of algorithms (see ~\cite{bubnow}). See ~\cite{bubnow} for a precise definition and
+or EXP family of algorithms (see~\cite{bubnow}). See~\cite{bubnow} for a precise definition and
 thorough treatment of this subject. These algorithms however all rely on
@@
 Table~\ref{tab:logs} presents the 7 logs, which can all be obtained from the
-Parallel Workload Archive ~\cite{Feitelson20142967}. We use the
-``cleaned'' ~\cite{pwa-clean} version of the logs as per the archive. We impose
+Parallel Workload Archive~\cite{Feitelson20142967}. We use the
+``cleaned''~\cite{pwa-clean} version of the logs as per the archive. We impose
 an additional filtering step to the workloads\footnote{This filtering step is
-available in the reproducible workflow ~\cite{repro} as shell script
+available in the reproducible workflow~\cite{repro} as shell script
 \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter|} in order to
@@
 This is the approach used by high-fidelity batch scheduling simulators such as
-BatSim ~\cite{batsim}. In the present work, the focus is on studying the
+BatSim~\cite{batsim}. In the present work, the focus is on studying the
 EASY-Backfilling mechanism in itself, without addressing the allocation
@@
 any available processor on the system. During this work, we developed a
-lightweight simulator ~\cite{ocst} and now release it to the community along with
+lightweight simulator~\cite{ocst} and now release it to the community along with
 this article. See the reproducibility paragraph below for more information.
@@
 Comparing algorithms in batch scheduling is made especially difficult by the
-fluctuations in commonly used metrics ~\cite{feitelson1998metrics}. The variability in
+fluctuations in commonly used metrics~\cite{feitelson1998metrics}. The variability in
 performance is known to outrank the available sample sizes: \textit{It is not
@@
 conclusions about which algorithm is better. This is a known
-problem ~\cite{flurries} that motivated decades of research in workload
-modeling ~\cite{feitbook}. Workload models have drawbacks in that they usually
+problem~\cite{flurries} that motivated decades of research in workload
+modeling~\cite{feitbook}. Workload models have drawbacks in that they usually
 lose the details of the workloads. This is problematic due to the non-linear
 aspect of the job scheduling process. This is discussed
-in ~\cite{feitresampling}, an approach that attempts to solve this problem by
+in~\cite{feitresampling}, an approach that attempts to solve this problem by
 resampling data directly from the original traces. In this work, we use a
@@
 threshold $\Theta$ at 40 hours. The choice of this default
-stems from the analysis developed in ~\cite{jsspp17}. Essentially, this value is high
+stems from the analysis developed in~\cite{jsspp17}. Essentially, this value is high
 enough for queue reorderings to be leveraged while being low enough to provide
@@
 
-We publish a declarative workflow ~\cite{repro} based on our
-simulator ~\cite{ocst} that runs experiments and generates all the figures from
+We publish a declarative workflow~\cite{repro} based on our
+simulator~\cite{ocst} that runs experiments and generates all the figures from
 this paper.
@@
 packaging software. As our experiments are CPU bound, we decide to opt for the
-software packaging approach ~\cite{nix}. This allows to replicate the experiments on
+software packaging approach~\cite{nix}. This allows to replicate the experiments on
 clusters where virtualization is not available or kernels are too old for
@@
 
-All the code associated with this work is released under the ISC ~\cite{isc}
+All the code associated with this work is released under the ISC~\cite{isc}
 license.
@@
 All data from this paper were obtained from the Swf Parallel Workloads
-Archive ~\cite{Feitelson20142967}.
-
- All experiments are tied together using zymake ~\cite{zymake}, a minimalistic
+Archive~\cite{Feitelson20142967}.
+
+ All experiments are tied together using zymake~\cite{zymake}, a minimalistic
  workflow system designed for computational experiments\footnote{The zymake system
@@
      simulator specially written for this work. This simulator is made
-     available under the ISC ~\cite{isc} license as a persistent zenodo
-     archive at ~\cite{ocst}.
+     available under the ISC~\cite{isc} license as a persistent zenodo
+     archive at~\cite{ocst}.
 

### 2018-01-12 paper/paper.tex
fixes

@@
 
-Let us emphasize the main results of this study: While the First-Come,
-First-Serve approach to minimizing the waiting times is a common default
-strategy, it can definitively be outperformed. We show how to automate selection
-of a good alternative policy in two cases (depending on whether a simulator is
-available or not), and provide the difference in performance to be expected
-between both situations.
+Let us emphasize the main results of this study. We show how to select a good
+alternative to the First-Come, First-Serve policy in an online manner. We
+present two automated methods for selecting a good alternative policy in two
+cases (depending on whether a simulator is available or not), and provide the
+difference in performance to be expected between both situations.
 
@@
 are stationary but the reader has to bear in mind that our framework can be use
-on different distributions by resorting to different values of lambda that can
+on different distributions by resorting to different values of $\lambda$ that can
 be set by cross-validation.

### 2018-01-12 paper/revision.tex
fixes

@@
 
-\title{Major revision comments.}
+\title{Comments on the revised version}
 
@@
 \maketitle
+
+Let us first sincerely thank the reviewers for their relevant and in-depth
+remarks. Their comments helped us to clarify the focus of this work, which
+lead to restructure the evaluation procedure to some extent. As part of these
+modifications, we changed the title to: "Online Tuning of EASY-Backfilling
+using Queue Reordering Policies". It better reflects the main goal, which is
+not to advocate bandit strategies, but rather to study their feasibility, and
+to show where the tradeoff is compared to using simulations for tuning the
+algorithms. We hope this revised version may now be considered for publication. 
+We detail our answers to rewiewer's comments (bold).
 
@@
 
+General comments: Our scope in this work is is not to advocate particular scheduling strategies (Bandit), but rather to study their feasibility, 
+and to show where the tradeoff is compared to using simulations for tuning the algorithms. 
+We are aware that is it not so common in our field where the classical approach is to formalize an hypothesis and to assess it by experiments. 
+The methodology that we propose here is to consider the 12 most popular (and efficient) reordering policies and to study their behavior.
+In the mean time, this also leads to figure out at the end if we can select the best ones.
+This methodology is closer to an empiristic study that are common in other fields.
+We added a brief justification at the end of the introduction.
+The reason why we have chosen this way was because it is very difficult to identify a priori what is the best policy and assess it. 
+Our attempt was to consider all possible policies. 
+\medskip
+
+We answer below more specific points:
+
+We focus in this paper on average waiting time. We have considered other objectives like bounded slow-down.
+This choice simply reflects the view of the users. Other metrics are indeed possible,
+and our method is agnostic to this choice.
+
+We agree that thresholding jobs is a standard approach and added a reference 
+that explores this aspect.
+
 \textbf{Similarly, the authors simply use 40h as threshold and simply state
@@
 
-This is indeed a mistake on our part which made the way to the final version,
-this is now corrected. The article is "Tuning Backfilling Queues" from JSSPP17.\\
+Right, this comes from a previous version, thanks for pointing this out. 
+We completed the reference: "Tuning Backfilling Queues" JSSPP17.
+\medskip
 
@@
 
+Right, we added an explanation for this in Section 6.1.3, page 8. This is due to the fact
+that the resampling technique generates stationary data.\\
+
 \textbf{Altogether, the reviewer believes that an experimental study with so
 many design problems is not relevant to the readers of TDPS. }
-
 \textbf{The reviewer does not see how the authors can overcome these
@@
 manuscript.}
+
+We disagree. We strongly believe that such a unusual approach is valuable,
+in the sense that it allows to get a sense of how useful simulators really are
+for the purpose of tuning backfilling heuristics.\\
 
@@
 
+Right, we polished the abstract accordingly\\
+
 \textbf{Previous publications explain that studies using traces suffer from
@@
 authors mention this problem but do not discuss their impact and strategies to
-avoid these problems although other publications present such strategies.
-Previous studies have described that the KTH trace is hardly comparable with
+avoid these problems although other publications present such strategies.}
+
+The use of our resampling method means the utilization is slightly lowered in
+the experiments comparing to the original traces. We do not believe this to be
+an issue, since we are studying adaptive methods in themselves, and not the 
+underlying systems.\\
+
+\textbf{Previous studies have described that the KTH trace is hardly comparable with
 other traces due to a specific acceptance policy. The authors do not discuss
-this problem. The authors use either 1 week or 1 day feedback justifying their
+this problem.}
+
+The traces we selected are quite disparate. This is on purpose to demonstrate
+the value of an adaptive approach: it works well on any kind of system.
+Moreover, our simulation model is simple: on real machines, we could reasonably
+expect the difference in behavior to increase. (See the answer to reviewer 3
+on this point.)\\
+
+\textbf{The authors use either 1 week or 1 day feedback justifying their
 choices with the periodicity of the traces. Although such approach is
@@
 periods as well and comparing the results. }
+
+We didn't set out to optimize the $\Delta$ hyperparameter. While we agree it
+could be optimized further similarly to $\epsilon$ by leave-one-out
+cross-validation, we don't believe it would add much to the discussion.\\
 
@@
 
+Right. We modified the introduction to better reflect the scope of the study.
+(See the last paragraph)\\
+
 \textbf{The authors can improve readability of the manuscript by explaining the
@@
 
-We added two figures to illustrate the algorithms.\\
+We added two figures to illustrate the algorithms in page 7 and 8.\\
 
@@
 reviewer does not understand why readers must contact the authors by email to
-obtain data to reproduce the study.  }
+obtain data to reproduce the study.}
 
 As previously concerning the missing reference, this was inadvertent on our
-part! This is fixed in the major revision. Apologies for this error.
+part! This is fixed in the major revision. 
+Thanks.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 following comments make some suggestions for possible improvements.}
+
+Thanks!
 
@@
 the title, to "Online Tuning of EASY-Backfilling using Queue Reordering
-Policies."\\
+Policies."
+\medskip
 
@@
 \end{enumerate}
-Thank you for your suggestions, we had not come across these papers! There are indeed very related to our effort.
+Thank you for your suggestions. These papers are indeed very related to our study. 
+They have been included in the references.
 
@@
 
-We did add a basic description of the setting in the description.\\
+Right, we clarified this by adding a sentence describing principles of the algorithms.
+\medskip
 
@@
 processors.  Similarly, in Section 4.1 I for one would find the names more
-intuitive with SJF/LJF instead of SPF/LPF (shortest/longes job first) SPF/LPF
+intuitive with SJF/LJF instead of SPF/LPF (shortest/longest job first) SPF/LPF
 instead of SQF/LQF (smallest/largest processors first) expansion factor is
@@
 
-We agree that $p_i$ is not very intuitive, but for lack of a standard we chose
-to use notation from scheduling theory.  In face of the many possible acronyms,
-we chose to rather be consistent with the $p_i$ and $q_i$ notation - SPF ist
-smallest $p_i$ first and SQF is smallest $q_i$ first..  We added the remark
+Well, our notations are coherent within the whole paper.
+We agree that $p_i$ may be confusing for some readers.
+However, since we lack of a standard, we chose
+to use the classical notation of scheduling theory. 
+Regarding the way to name the policies, they are different acronyms used in the literature,
+again we chose to be consistent with the $p_i$ and $q_i$ notations - SPF corresponds to the
+smallest $p_i$ first and SQF refers to the smallest $q_i$ first.  We added the remark
 concerning the total resources and the expansion factor. This notation comes
 from the "Characterization of backfilling strategies for parallel job
-scheduling" article, where the authors call Expansion Factor the value of the
-slowdown where the upper bound on the running-time is used.\\
+scheduling" paper, where the authors call \textit{Expansion Factor} the value of the
+slowdown where the upper bound on the running-time is used. 
+\medskip
 
@@
 makes the study more realistic with respect to the cases where simulations are
-expensive(which is most large systems).  Indeed, the more realistic simulations
-are, the more time they take. For example, precise simulation frameworks such
-as SimGrid/Batsim have high runtimes, and simulating the history of the whole
-system every period is prohibitive if the network topology and/or multiple
+expensive (which is the case for most large systems).  Indeed, the more realistic simulations, 
+the larger time they take. For example, precise simulation frameworks such
+as SimGrid/Batsim have large runtimes, and simulating the history of the whole
+system every period is too prohibitive if the network topology and/or multiple
 resource constraints are used. As for discounting - it can indeed be used
-in both cases.\\
+in both cases.
+\medskip
 
@@
 
-We do agree that this does not change the estimate - the point of adding 
+We agree that this does not change the estimate - the point of adding 
 this noise was to show that the sample size is not prohibitive in practice.
 Indeed, we learn the policy almost as fast with this noise, which
-motivates the use of the technique in real-world cases. The point of this noise
-is really to compare the difference in the shapes of the two green curves
-in Figure 3.\\
+motivates the use of the technique in real-world cases. The main motivation for introducing this noise
+was really to compare the difference in the shapes of the two green curves
+in Figure 3.
+\medskip
 
@@
 
-You are right - and we made this point clear in the text.\\
+You are right, we clarified this point in the text (SECTION xx - PAGE xx).\\
 
@@
 
-Doing this would drown the point in the author's opinion. We agree it is an
-interesting point - we now refer to our previous publication 'Tuning
-Backfilling Queues'(JSSPP17), which was inadvertently ommited from the related
-works of the first submission.\\
+Doing this would drown the point in the author's opinion. 
+QUE VEUX-TU DIRE VRAIMENT ICI?
+Good remark, we agree it is an interesting point, which was addressed in our previous publication 'Tuning
+Backfilling Queues' (JSSPP17).
+\medskip
 
-  \textbf{Section 6.3.2 FCFS vs. Random: you justify the result be a conjecture that jobs do not pile up in the queue.  you should verify this by counting the jobs int he queue in the simulation.}
+  \textbf{Section 6.3.2 FCFS vs. Random: you justify the result be a conjecture that jobs do not pile up in the queue.  
+  You should verify this by counting the jobs int he queue in the simulation.}
 
-We aggree that this conjecture is not warranted - we removed it in the revision.
-We do have plans to explore this aspect for a future publication where the
-discussion will be more appropriate.\\
+We agree that this conjecture is not warranted, and thus, we removed it in the revision.
+We have plans to explore this aspect in a future work where the
+discussion would be more appropriate.
+\medskip
 
@@
 
-Done, thanks you for the suggestion.\\
+Done, thanks you for the suggestion.
+\medskip
 
@@
 decaying exploration does provide better performance, it is also harder to
-tune.\\
+tune.
+\medskip
 
@@
 
-This is corrected in the major revision - we inadvertently ommited this
-when preparing the submission. The reference is 'Tuning Backfilling Queues'(JSSPP17).\\
+Right, this is corrected in the major revision. The reference is 'Tuning Backfilling Queues' (JSSPP17)
+and it has been added in the references.
+\medskip
 
@@
 
-We incorporated these changes and thank the reviewer for their detailed read.\\
+We incorporated all the suggested changes, thanks.
+\medskip
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-We do agree with the reviewer that in order to really motivate the use of
-bandit/resimulation aproaches, one should compare the performance of static
+Right, we agree that in order to really motivate the use of
+bandit/resimulation approaches, one should compare the performance of static
 policies, say using leave-one-trace-out cross validation, to the performance of
@@
 hyperparameters.
-However, this work does not aim to underline the necessity of the bandit
-approach (the title has been changed and the introduction modified in the major
-revision to reflect this better). Rather, we try to show its advantages,
+However, this work does not aim to emphasize the necessity of the bandit
+approach (the title has been changed and the introduction has been modified in the 
+revision to better reflect this). Our goal was rather to show its advantages,
 feasibility and limitations. 
 There are two main reasons for not trying to be too demonstrative of the superiority
-of the adaptive approach using our hypotheses.
+of the adaptive approach using our hypotheses, namely:
 
-First, this study uses a limited amount of workload logs. In consequence, the
+(1) First, this study used a limited amount of workload logs. In consequence, the
 process of comparing the adaptive (hyperparameter-cross-validated) choice to
 the static (cross-validated in the set of fixed policies) choice would not be
-really conclusive.  Indeed, the more workload logs are presented, the more
-diversity is present and the more adaptation is necessary.
-Second, we use a simplistic simulation model. In this model, we do make
+really conclusive.  Indeed, the more the number of workload logs, the higher the
+diversity and the more adaptations are needed.
+
+(2) Second, we used a light simulation model. In this model, we do make
 machines *more* similar to each other by neglecting topological differences. It
@@
 We modified the present experiments to use leave-one-out cross
-validation in the hyperparameter selection, and made the reasons for this
-choice of experimental protocol clearer in the text.
+validation in the hyperparameter selection, and we clarified the reasons for
+chosing this experimental protocol.
 We are preparing a subsequent study using an approach closer to what you
-describe, in which we greatly increase the number of workload logs used.\\
+describe, in which we greatly increase the number of workload logs used.
+\medskip
 
-\textbf{The authors assume +/-20% noise in the simulation running in the
+\textbf{The authors assume +/-20 \% noise in the simulation running in the
 proposed strategy. However, the reason why this assumption is suitable is not
-presented. The authors should discuss the noise of 20% is suitable to represent
+presented. The authors should discuss the noise of 20 \% is suitable to represent
 real world jobs.}
 
-A justification was indeed missing, this value is taken as to be sufficiently
-larger than values previously reported, see e.g. Batsim: a Realistic
-Language-Independent Resources and Jobs Management Systems Simulator which
-reports 2 percent variations of the simulated metric compared to the metric on
-the real systems. We added this remark to the text.\\
+Right, a justification was clearly missing, this value is taken as to be
+sufficiently larger than the values previously reported, see e.g. "Batsim: a
+Realistic Language-Independent Resources and Jobs Management Systems Simulator"
+which reports few percents of variation of the simulated metric compared to the
+metric on the real systems. We added this remark in the text.\\
 
@@
 
-We incorporated these changes and thank the reviewer for their detailed read.\\
+Thanks, we incorporated all the corresponding changes in the text.
 

### 2018-03-02 paper/revision2.tex
revision2

\documentclass[]{article}

\usepackage{listings}
\usepackage{fullpage}
\usepackage{dirtree}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bbm}
\usepackage[scientific-notation=true]{siunitx}

\usepackage{amssymb}
\usepackage{tabulary}

\usepackage{etoolbox}

\usepackage{algorithm}
\usepackage{algorithmic}

\title{Comments on the revised version}

\begin{document}

\maketitle

Let us first sincerely thank the reviewers for their relevant and in-depth
remarks. Their comments helped us to clarify the focus of this work, which
lead to restructure the evaluation procedure to some extent. As part of these
modifications, we changed the title to: "Online Tuning of EASY-Backfilling
using Queue Reordering Policies". It better reflects the main goal, which is
not to advocate bandit strategies, but rather to study their feasibility, and
to show where the tradeoff is compared to using simulations for tuning the
algorithms. We hope this revised version may now be considered for publication. 
We detail our answers to rewiewer's comments (bold).

\section{Reviewer 1}

\textbf{
  I basically liked this paper and support its publication.  I have only very few and minor comments.}

\textbf{In sect 1.1 you claim 2 "new" online approaches, and in 1.2 you call them "original". Your work is valid even if you
admit that both have been suggested before.
}

\textbf{Last paragraph of 1.2 is the third repetition of this.  It is redundant, despite the fact that some reviewer didn't get
it.
}

\textbf{In sect 5.1 you explain that you set lambda=1 because due to the resampling the distributions are stationary.  That is
correct.  But it would have been good to also conduct simulations on the original traces, which are probably not
stationary, and therefore there is more to gain from switching between policies.
Also, instead of starting the simulation of each period from an empty system, why not start from the situation at the
end of the previous period?
}

\textbf{Typo at end of 6.1.3: sentence with [24] appears twice.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Reviewer 2}

\textbf{In comparison to the first version, the authors have made several small changes that results in minor improvements of
the manuscript. Nevertheless, the main problems of the manuscript remain. Therefore, the position of the reviewer
remains unchanged.
}

\textbf{The authors present a study based on a selection of several strategies and parameters. They execute several simulations
and show that using a feedback mechanism improvements over the standard approach without feedback are possible. This
results is not surprising considering previous results some of which are referenced within the manuscript. It must be
noted that the authors do not suggest a specific new scheduling algorithm and show that this algorithm results in
significant improvements over existing methods.
}

\textbf{The authors mention "reducing uncertainty" in Section 1.1. It is not clear why this reduction is relevant within the
manuscript (except as a justification for adding noise?)
The first paragraph of Section 1.1 is relevant to the content of the manuscript. The study of the manuscript uses old
traces that are far from extreme scale parallel computing. The relevance to distributed computing is not clear since
only traces from parallel computers are used.
}

\textbf{Why do the authors discuss conservative backfilling in Section 2 although it does not matter in the rest of the
manuscript?
What is the meaning of "dependency of scheduling metrics on the workload"? The authors only apply a single scheduling
metric (average waiting time) and do not discuss such dependency. Why is it important that previous work does not
address this dependency in the context of this manuscript?
}

\textbf{The authors mention duality between cumulative and maximal scheduling costs. Why is this duality relevant in the
context of this manuscript? In Section 3.3 the authors claim that "Section 4.2 will outline our approach to address the
biobjective aspect of this problem". However, Section 4.2. is very brief and does not discuss such aspect. The reviewer
did not find any discussion in the rest of the manuscript. Altogether not all references cited in the manuscript are
relevant for the content of the manuscript.
}

\textbf{The design of the study is contradictory: on the one hand, the authors only select a single metric since this metric
"is one of the more commonly used objectives". On the other hand, they select 12 reordering strategies without giving
any justification based on the strategy. They simply state that they want the search space "being as semantically
diverse as possible" without stating that they achieved this goal with the selection of these strategies. Why do they
need all the 12 strategies and why not using more strategies? Again they only remark that the strategies "include most
policies from related works that we are aware of." Such justification is hardly sufficient for a scientific study.
}

\textbf{In Section 5, the authors claim that "the distribution of the job submission process does not radically differ between
consecutive periods." This claim is not true for every period length. For instance, when using 12 hours as period
length, two consecutive periods may cover day and night with different submission processes.
}

\textbf{The allocation of jobs to periods based on the submission time may be problematic. For instance a long running job may
be submitted in one period, waiting for execution in the next period and then executing in the next several periods.
Therefore, it influences several periods. The authors briefly mention such dependencies (Section 5.3) and state that
there may be a bias without discussing this issue.
}

\textbf{The authors still keep the decaying discount factor $\lambda$ without using it in the study. In an exploratory study, the
authors can either ignore such factor because it is not relevant (and saying why this claim holds) or they must include
it into the study.
}

\textbf{The authors introduce a noise factor to deal with imprecision. This way they introduce an approach to model
imprecision. But they do not show that their model is suitable in reality.
}

\textbf{In Section 5.2, the authors claim that the length of the period has little impact here as the cost metric corresponds
to the cumulative simulated waiting time for each policy. As already mentioned, this may not be true for periods that
are small compared to the waiting time of jobs.
}

\textbf{Figure 3 does not illustrate the process of the $\epsilon$-greedy strategy since there is no $\epsilon$ in Figure 3.
}

\textbf{Section 6.1.1 mentions cores in traces although several of the old traces use machines that do not have multicore
processors.
}

\textbf{The authors use 40 hours as threshold and justify this selection by referring to another publication. The reader does
not know about the arguments without reading this other publication. Since the design of the study is at the core of
this manuscript, such justification if not acceptable.
}

\textbf{The authors state that they have simulated a total of 92400 years. However, this absolute figure is not relevant
without showing the coverage of the problem space.
}

\textbf{There are only few language problems in the manuscript:
Software and work are used in the singular form. The authors must adjust the verb correspondingly (Section 1.1)
wokload in Fig. 2 and 3
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Reviewer 3}

\textbf{I think that the simulation model presented in this paper is carefully created and it shows convincing results.
However, there are two questions about the discussion for the simulation results:
}

\textbf{The results in Table 3 and Table 4 show that the performance of the static strategy, LQF, is comparable with the
presented online tuning strategies, Full and Bandit. Which strategy is better in the real world? We can say that LQF
may be preferable because they are less complicated and run with smaller scheduling overhead. The authors should add
the discussion to compare LQF and the online tuning strategies.
}

\textbf{The authors conclude that simulation-based strategy is best but the bandit-based strategy is easier to use and cheaper
to run. How easy and how cheaper can we run the bandit-based strategy? The authors should add the quantitative
discussion how the bandit-based strategy is easy to use and saves cost (or scheduling overhead) compared with the
simulation-based strategies.
}

\bibliographystyle{abbrv}
\bibliography{bibliography}

\end{document}

### 2018-03-02 paper/paper.tex

@@
 attainable performance in the search set defined in~\ref{subsec:policies}.
-Note that there seems to be some regularity; for instance the SAF policy is remarkably
-inefficient and the LQF policy works very well. However there are traces for which LQF is not optimal. Moreover, these
-are~\textit{a-posteriori} results, which means that we only know
+Note that there seems to be some regularity; for instance the SAF policy is
+remarkably inefficient and the LQF policy works very well. However there are
+traces for which LQF is not optimal.  For these machines, ordering jobs by
+decreasing size seems to be more efficient than ordering them by time. This
+study is limited in scope, and these results are possibly different on other
+machines. Moreover, the simulator model is simplistic and the relative order of
+policies may change, for instance when introducing a topology model in the
+simulations. This is however sufficient to study adaptive policies.
+
+These are~\textit{a-posteriori} results, which means that we only know
 \textit{post-factum} that this policy would have been better.

### 2018-03-02 paper/revision.tex

@@
 
-You are right, we clarified this point in the text (SECTION xx - PAGE xx).\\
+You are right, we clarified this point in the text (Section 6.6.1 - Page 7).\\
 
@@
 
-Doing this would drown the point in the author's opinion. 
-QUE VEUX-TU DIRE VRAIMENT ICI?
-Good remark, we agree it is an interesting point, which was addressed in our previous publication 'Tuning
-Backfilling Queues' (JSSPP17).
+It is an interesting point, which was addressed in our previous publication
+'Tuning Backfilling Queues' (JSSPP17). Here, we prefer to be cautious. Indeed,
+the high number of simulations in this work forced us to use a simplistic
+simulator model. We can reasonably expect these results to change when adding
+a topology model, or a model of job interference for example. While we believe
+this simple model is sufficient to study the behavior of adaptive policies, we
+prefer not to draw conclusions about the systems themselves from these
+simulations. We added this remark in the text (Section 6.3.1 - page 10).
+
 \medskip
@@
 
-(2) Second, we used a light simulation model. In this model, we do make
+(2) Second, we need to run high number of simulations, and accordingly we used a light simulation model . In this model, we do make
 machines *more* similar to each other by neglecting topological differences. It
@@
 rather than decrease, when moving away from a simplistic model to the real
-systems.
+systems. This reason is also why we don't expand on which fixed strategy is 
+the best. Indeed, the results are probably different when adding a topology model,
+job interference, or an I/O congestion model. We do believe this simple
+model is enough for studying adaptive policies though. We added a remark in 
+the text (Section 6.3.1 - Page 10)
+
 We modified the present experiments to use leave-one-out cross
@@
 which reports few percents of variation of the simulated metric compared to the
-metric on the real systems. We added this remark in the text.\\
+metric on the real systems. We added this remark in the text (Section 5.2- Page 6).\\
 

### 2018-03-07 paper/revision2.tex
zmk fix for the S/L inversion

@@
 }
+You are right. We rephrased this in a more neutral way.
 
@@
 }
+Regarding your first point, good remark, we based this experimental design around previous results
+from "Tuning Backfilling Queues"(JSSPP17) that show some degree of stationarity. We indeed
+expect that non-stationary policy adaptation may allow more improvements (while
+being harder to achieve). We made this clearer in the text.
+
+Your second point is also relevant, yes that should perhaps slightly improve the results.
+But we can see that the 'full feedback' simulated policy here is almost as good as the best
+policy in the set, so we did not worry about this. In a non-stationary case, it may indeed
+be necessary to properly take care of this.
 
@@
 }
+Right, we fixed this.
 
@@
 }
+We regret that the detailed answer of the previous rebuttal did not convince you. We disagree with this statement. We indeed
+provide a new scheduling algorithm, which adapts to the situation by selecting the correct reordering policy. This algorithm
+results in significant improvements over existing methods.
+Thank you for your in-depth remarks once again. We integrated these into the minor revision.
 
@@
 
+First of all, we corrected a crucial typographical error in the text. The S and
+L were inverted in both Table 3 and Figure 4. This resolves a blatant
+consistency issue with other publications, most crucially "Tuning Backfilling
+Queues"(JSSPP17).
+
+Indeed, the SQF policy works quite well on the traces considered here on average, and
+under this resampling scheme. The variability however, is high, and it is not the best
+on other traces.
+It would indeed be interesting to have an experimental evaluation of exactly how much
+one gains on average by choosing the bandit over a leave-one-trace-out majority vote policy
+choice.
+
+However, a good estimation of this quantity requires using more traces, and perhaps factoring
+in the topology model of the machines, which should increase the heterogeneity between platforms
+and therefore increase the dependence of the simulation performance on the trace/platform.
+Such a study would be much more tedious.
+Here, we prove the feasibility of the approach. The need for adaptivity is actually understated
+by our simulation approach.
+
 \textbf{The authors conclude that simulation-based strategy is best but the bandit-based strategy is easier to use and cheaper

### 2018-03-07 paper/paper.tex
updates

@@
 so-called EASY\footnote{Extensible Argonne Scheduling sYstem
-(EASY)}-Backfilling family of heuristics using two new online approaches. One
+(EASY)}-Backfilling family of heuristics using two online approaches. One
 approach is based on a simulator and the second uses a multi-armed bandit
@@
 
-Our main contribution is to investigate two original methods for learning a
+Our main contribution is to investigate two methods for learning a
 good scheduling policy on a per-system basis. In this work, we focus on the
@@
 it is reasonable
-to assume that the system is in a stationary state.~\cite{jsspp17} and assume for these experiments that the system is in a
-stationary state. We do provide guidelines as how to incorporate tracking
+to assume that the system is in a stationary state. Indeed,~\cite{jsspp17} performs
+train/test experiments using the same policy space as we use here. We do provide guidelines as how to incorporate tracking
 into our algorithms in Section~\ref{sec:online} by introducing a decaying
@@
     \hline
-    Trace        & LCFS  & LPF            & SPF   & LQF            & SQF   & SEXP  & LEXP           & LRF            & SRF   & LAF            & SAF   \\
+    Trace        & LCFS  & SPF            & LPF   & SQF            & LQF   & LEXP  & SEXP           & SRF            & LRF   & SAF            & LAF   \\
     \hline
@@
     \hline
-    Trace        & Best(hindsight) & Random &       & Full           &                & Noisy &       & Bandit & \\
+    Trace        & Best(hindsight) & & Random       &            &Full                & & Noisy        & & Bandit  \\
     \hline
     $\Delta$     &        & week   & day   & week           & day            & week  & day   & week   & day \\
-    KTH-SP2      & -16\%  & -6\%   & -8\%  & \textbf{-12}\% & -\textbf{11}\% & -12\% & -12\% & -7\%   & -10\% \\
+    KTH-SP2      & -16\%  & -6\%   & -8\%  & \textbf{-12}\% & 11\% & -\textbf{-12}\% & -\textbf{-12}\% & -7\%   & -10\% \\
     CTC-SP2      & -47\%  & -20\%  & -23\% & \textbf{-44}\% & \textbf{-44}\% & -44\% & -44\% & -26\%  & -31\% \\
-    SDSC-SP2     & -15\%  & -4\%   & -7\%  & \textbf{-13}\% & \textbf{-13}\% & -13\% & -13\% & -5\%   & -8\% \\
+    SDSC-SP2     & -15\%  & -4\%   & -7\%  & \textbf{-13}\% & \textbf{-13}\% & -\textbf{-13}\% & -\textbf{-13}\% & -5\%   & -8\% \\
     SDSC-BLUE    & -32\%  & -12\%  & -14\% & \textbf{-31}\% & \textbf{-31}\% & -30\% & -29\% & -12\%  & -17\% \\
     ANL-Intrepid & -28\%  & -13\%  & -20\% & -22\%          & \textbf{-26}\% & -19\% & -24\% & -13\%  & -20\% \\
-    CEA-Curie    & -57\%  & -35\%  & -42\% & \textbf{-53}\% & -47\%          & -53\% & -48\% & -36\%  & -46\% \\
+    CEA-Curie    & -57\%  & -35\%  & -42\% & \textbf{-53}\% & -47\%          & -\textbf{-53}\% & -48\% & -36\%  & -46\% \\
     Unilu-Gaia   & -62\%  & -29\%  & -31\% & -59\%          & \textbf{-60}\% & -58\% & -58\% & -33\%  & -34\% \\
@@
 
-There are two directions to extend this work. First, the question of
-the existence of a more efficient selection strategy for bandit feedback for
-these systems is not addressed by this work. Is a simulator really necessary?
-%@Denis: a essayer.. 1 mois de travail pour tester?
-Second, one can't help but wonder what performance could be achieved by extending
-the search space to a wider class of policies. This means studying learning performance using some
-function approximator.
-%@Denis: c'est ce que font vasily et salah de deux manieres differentes.
-%TODO SPLIT IN TWO, MENTION FACTOR ANALYSIS.
+There are three directions to extend this work. First, the question of the
+existence of a more efficient selection strategy for bandit feedback for these
+systems is not addressed by this work. Is a simulator really necessary?
+Second, one can't help but wonder what performance could be achieved by
+extending the search space to a wider class of policies. This means studying
+learning performance using some function approximator.
+
 As both these directions are data-bound, they will require careful application of

### 2018-03-07 paper/revision2.tex
updates

@@
 Let us first sincerely thank the reviewers for their relevant and in-depth
-remarks. Their comments helped us to clarify the focus of this work, which
-lead to restructure the evaluation procedure to some extent. As part of these
-modifications, we changed the title to: "Online Tuning of EASY-Backfilling
-using Queue Reordering Policies". It better reflects the main goal, which is
-not to advocate bandit strategies, but rather to study their feasibility, and
-to show where the tradeoff is compared to using simulations for tuning the
-algorithms. We hope this revised version may now be considered for publication. 
+remarks once again. Their comments helped us further the discussion of the
+results, and find a typographical error that passed our scrutiny.
+
 We detail our answers to rewiewer's comments (bold).
@@
 }
+
 You are right. We rephrased this in a more neutral way.
@@
 }
-Regarding your first point, good remark, we based this experimental design around previous results
-from "Tuning Backfilling Queues"(JSSPP17) that show some degree of stationarity. We indeed
-expect that non-stationary policy adaptation may allow more improvements (while
-being harder to achieve). We made this clearer in the text.
 
-Your second point is also relevant, yes that should perhaps slightly improve the results.
+Regarding your first point, good remark, we based this experimental design
+around previous results from "Tuning Backfilling Queues"(JSSPP17) that show
+some degree of stationarity using a train/test procedure. We made this clearer
+in the text.
+We indeed expect that non-stationary policy adaptation may allow more
+improvements (while simultaneously being harder to achieve).
+
+Your second point is also relevant, yes that should improve the results.
 But we can see that the 'full feedback' simulated policy here is almost as good as the best
@@
 }
+
 Right, we fixed this.
@@
 }
+
 We regret that the detailed answer of the previous rebuttal did not convince you. We disagree with this statement. We indeed
@@
 results in significant improvements over existing methods.
-Thank you for your in-depth remarks once again. We integrated these into the minor revision.
+We thank you for your in-depth remarks once again. We integrated these comments into the minor revision.
 
@@
 
-First of all, we corrected a crucial typographical error in the text. The S and
+Right, this discussion is warranted. We added paragraph X.
+
+First of all, we corrected a crucial typographical error in the text that slipped
+our scrutiny at the time of running the build chain for the publication. The S and
 L were inverted in both Table 3 and Figure 4. This resolves a blatant
@@
 
-Indeed, the SQF policy works quite well on the traces considered here on average, and
-under this resampling scheme. The variability however, is high, and it is not the best
-on other traces.
-It would indeed be interesting to have an experimental evaluation of exactly how much
-one gains on average by choosing the bandit over a leave-one-trace-out majority vote policy
-choice.
+Indeed, the SQF policy works quite well on the traces considered here on
+average, and under this resampling scheme. The variability however, is high,
+and it is not the best on other traces.  It would indeed be interesting to have
+an experimental evaluation of exactly how much one gains on average by choosing
+the bandit over a leave-one-trace-out majority vote policy choice.
 
-However, a good estimation of this quantity requires using more traces, and perhaps factoring
-in the topology model of the machines, which should increase the heterogeneity between platforms
-and therefore increase the dependence of the simulation performance on the trace/platform.
-Such a study would be much more tedious.
-Here, we prove the feasibility of the approach. The need for adaptivity is actually understated
-by our simulation approach.
+However, a good estimation of this quantity requires using more traces, and
+perhaps factoring in the topology model of the machines, which should increase
+the heterogeneity between platforms and therefore increase the dependence of
+the simulation performance on the trace/platform. Such a study would be much
+more tedious. Here, we prove the feasibility of the approach. The need for
+adaptivity is actually understated by our simulation approach, but in all
+honesty we don't believe to be able to measure this rigorously.
 
@@
 
+Right, we need to make this clearer. We added this discussion to paragraph X.
+
+In both these cases, the scheduling decisions do not have an overhead, as the
+policy is fixed during the period (one day, or week).
+
+The bandit strategy has a negligible computational overhead at period end. Its two
+computational operations are maintaining a sequential average and generating a
+pseudo-random number.
+
+The simulation-based strategy requires to simulate the system K times every
+period end, where K is the number of alternative policies considered (here,
+K=12). Using our simulator, This takes a few seconds. Using a more precise
+simulator with topological modeling, one can argue that the overhead is still
+manageable since these simulations can be easily done in parallel. The most
+important constraint is that these simulations should not take too much time
+compared to the period size, as to not delay the policy switch decision too
+much.
+
 \bibliographystyle{abbrv}

### 2018-03-08 paper/paper.tex
revision2 first draft.

@@
 study more sophisticated policies that could leverage these data. During the
-last few years, there was an explosion in the number of works at the interface
+last few years, there was an explosion in the amount of work at the interface
 of the HPC and Big Data fields, dealing with learning algorithms. Recent
@@
 the distribution of the job submission process does not radically differ
-between consecutive periods. This assumption is validated in the study
+between consecutive weekly periods. This assumption is validated in the study
 presented in~\cite{jsspp17}. It further entails that the behavior of a policy
@@
 distributions where $\lambda$ can be set to 1 to non-stationary distributions
-where the more recent history has to be privileged. As explained in Section
-6.1, in our experiments $\lambda$ is set to 1 as the distributions considered
-are stationary but the reader has to bear in mind that our framework can be use
-on different distributions by resorting to different values of $\lambda$ that can
-be set by cross-validation.
+where the more recent history has to be privileged. As explained in
+Section~\ref{ssub:resampling}, in our experiments $\lambda$ is set to 1 as the
+distributions considered are stationary but the reader has to bear in mind that
+our framework can be use on different distributions by resorting to different
+values of $\lambda$ that can be set by cross-validation.
 
@@
   \caption{Illustration of the online simulation-based strategy. Every week,
-  the wokload from the past week is used to run simulations using every
+  the workload from the past week is used to run simulations using every
   reordering policy. The "model" $w$ is updated according
@@
 
-  \caption{Illustration of the bandit-based strategy. Every week,
+  \caption{Illustration of the bandit-based setting. Every week,
   the wokload from the past week is used to run simulations using every
@@
 Comparing algorithms in batch scheduling is made especially difficult by the
-fluctuations in commonly used metrics~\cite{feitelson1998metrics}. The variability in
-performance is known to outrank the available sample sizes: \textit{It is not
-enough to simply replay an algorithm using a workload trace}. Indeed, as
-presented in Figure~\ref{fig:variability}, the objective function we are using
-in this work is subject to high variability. In our experience, randomly
-shuffling months in a trace at random is enough to generate contradictory
-conclusions about which algorithm is better. This is a known
+fluctuations in commonly used metrics~\cite{feitelson1998metrics}. The
+variability in performance is known to outrank the available sample sizes:
+\textit{It is not enough to simply replay an algorithm using a workload trace}.
+Indeed, as presented in Figure~\ref{fig:variability}, the objective function we
+are using in this work is subject to high variability. In our experience,
+randomly shuffling months in a trace at random is enough to generate
+contradictory conclusions about which algorithm is better. This is a known
 problem~\cite{flurries} that motivated decades of research in workload
@@
 us to preserve the dependency within a week. Although the system state may not
-be independent from one week to an other, as illustrated in~\cite{jsspp17},
-it is reasonable
-to assume that the system is in a stationary state. Indeed,~\cite{jsspp17} performs
-train/test experiments using the same policy space as we use here. We do provide guidelines as how to incorporate tracking
+be independent from one week to an other, as illustrated in~\cite{jsspp17}, it
+is reasonable to assume that the system is in a stationary state.
+Indeed,~\cite{jsspp17} performs train/test experiments using the same policy
+space as we use here. We do provide guidelines as how to incorporate tracking
 into our algorithms in Section~\ref{sec:online} by introducing a decaying
-discount factor $\lambda$. In our experiments, we use stationary resampled data,
-and therefore no decay is necessary: $\lambda$ here is set to $1$.
+discount factor $\lambda$. In our experiments, we use stationary resampled
+data, and therefore no decay is necessary: $\lambda$ here is set to $1$.
 
@@
 
+\textbf{SQF versus online tuning.}
+Table~\ref{tab:full} shows that the SQF policy
+works quite well on the traces considered here on average, and under this
+resampling scheme. The variability however, is high, and this policy is not the
+best on all traces. It interesting to have an experimental evaluation of
+exactly how much one gains on average by using an online policy versus a
+leave-one-trace-out majority vote policy choice.  However, a good estimation of
+this quantity is tricky. Indeed, we expect the change in the relative
+performance of policies to increase with two factors. First, simply adding
+workload traces to our selection will naturally show more cases where behavior
+is different. Second, factoring in the topology model of the machines should
+increase the heterogeneity between platforms and therefore increase the
+dependence of the simulation performance on the trace/platform. Such a study
+would be much more tedious. Here the need for adaptivity is actually
+understated by the simulation approach.
+
+While we can not conclude on how much one gains by not using SQF in this case,
+we can comment on the added complexity and computing overhead of using the
+simulation-based and bandit online strategies.  In both these cases, the
+scheduling decisions do not have an overhead, as the policy is fixed during the
+period (one day, or week). The bandit strategy has a negligible computational
+overhead at period end. Its two computational operations are maintaining a
+sequential average and generating a pseudo-random number.  The simulation-based
+strategy requires to simulate the system K times every period end, where K is
+the number of alternative policies considered (here, K=12). Using our
+simulator, This takes a few seconds. Using a more precise simulator with
+topological modeling, one can argue that the overhead is still manageable since
+these simulations can be easily done in parallel. The most important constraint
+is that these simulations should not take too much time compared to the period
+size, as to not delay the policy switch decision too much.
+
 \begin{figure}[]

### 2018-03-08 paper/revision2.tex
revision2 first draft.

@@
 
+\usepackage{csquotes}
 \usepackage{listings}
@@
 remarks once again. Their comments helped us further the discussion of the
-results, and find a typographical error that passed our scrutiny.
+results, and find a typographical error in the presentation of the results that
+passed our scrutiny.
 
@@
 You are right. We rephrased this in a more neutral way.
+
+\medskip
 
@@
 
+\medskip
+
 \textbf{Typo at end of 6.1.3: sentence with [24] appears twice.
@@
 Right, we fixed this.
+
+\medskip
 
@@
 
-We regret that the detailed answer of the previous rebuttal did not convince you. We disagree with this statement. We indeed
+We regret that the detailed answer to your comments did not convince you. We disagree with this statement. We indeed
 provide a new scheduling algorithm, which adapts to the situation by selecting the correct reordering policy. This algorithm
@@
 We thank you for your in-depth remarks once again. We integrated these comments into the minor revision.
+
+\medskip
 
@@
 
-Right, this discussion is warranted. We added paragraph X.
-
 First of all, we corrected a crucial typographical error in the text that slipped
@@
 
-Indeed, the SQF policy works quite well on the traces considered here on
-average, and under this resampling scheme. The variability however, is high,
-and it is not the best on other traces.  It would indeed be interesting to have
-an experimental evaluation of exactly how much one gains on average by choosing
-the bandit over a leave-one-trace-out majority vote policy choice.
+You are right, the discussion you suggest is warranted. We added it to
+paragraph \textbf{SQF versus online tuning} in Section 6.3.2:
 
-However, a good estimation of this quantity requires using more traces, and
-perhaps factoring in the topology model of the machines, which should increase
-the heterogeneity between platforms and therefore increase the dependence of
-the simulation performance on the trace/platform. Such a study would be much
-more tedious. Here, we prove the feasibility of the approach. The need for
-adaptivity is actually understated by our simulation approach, but in all
-honesty we don't believe to be able to measure this rigorously.
+\begin{displayquote}
+Table~3 shows that the SQF policy
+works quite well on the traces considered here on average, and under this
+resampling scheme. The variability however, is high, and this policy is not the
+best on all traces. It interesting to have an experimental evaluation of
+exactly how much one gains on average by using an online policy versus a
+leave-one-trace-out majority vote policy choice.  However, a good estimation of
+this quantity is tricky. Indeed, we expect the change in the relative
+performance of policies to increase with two factors. First, simply adding
+workload traces to our selection will naturally show more cases where behavior
+is different. Second, factoring in the topology model of the machines should
+increase the heterogeneity between platforms and therefore increase the
+dependence of the simulation performance on the trace/platform. Such a study
+would be much more tedious. Here the need for adaptivity is actually
+understated by the simulation approach.
+\end{displayquote}
+
+\medskip
 
@@
 
-Right, we need to make this clearer. We added this discussion to paragraph X.
+Right, we need to make this clearer. We added this discussion to paragraph
+\textbf{SQF versus online tuning} in Section 6.3.2:
 
-In both these cases, the scheduling decisions do not have an overhead, as the
-policy is fixed during the period (one day, or week).
+\begin{displayquote}
+While we can not conclude on how much one gains by not using SQF in this case,
+we can comment on the added complexity and computing overhead of using the
+simulation-based and bandit online strategies. In both these cases, the
+scheduling decisions do not have an overhead, as the policy is fixed during the
+period (one day, or week). The bandit strategy has a negligible computational
+overhead at period end. Its two computational operations are maintaining a
+sequential average and generating a pseudo-random number.  The simulation-based
+strategy requires to simulate the system K times every period end, where K is
+the number of alternative policies considered (here, K=12). Using our
+simulator, This takes a few seconds. Using a more precise simulator with
+topological modeling, one can argue that the overhead is still manageable since
+these simulations can be easily done in parallel. The most important constraint
+is that these simulations should not take too much time compared to the period
+size, as to not delay the policy switch decision too much.
+\end{displayquote}
 
-The bandit strategy has a negligible computational overhead at period end. Its two
-computational operations are maintaining a sequential average and generating a
-pseudo-random number.
-
-The simulation-based strategy requires to simulate the system K times every
-period end, where K is the number of alternative policies considered (here,
-K=12). Using our simulator, This takes a few seconds. Using a more precise
-simulator with topological modeling, one can argue that the overhead is still
-manageable since these simulations can be easily done in parallel. The most
-important constraint is that these simulations should not take too much time
-compared to the period size, as to not delay the policy switch decision too
-much.
+\medskip
 

### 2018-03-09 paper/paper.tex
fix

@@
 system and wait for the system to be empty when job submissions cease at the
-end of the period (see Figure~\ref{fig:resimu}).
+end of the period (see Figure~\ref{fig:resimu}). Note that starting the simulation
+using the system state at the end of the previous period should improve the results.
+Here it will not be necessary hence we keep the simple version with an empty system.
+However, it may prove useful in the non-stationary case mentionned above.
 
@@
 \textbf{SQF versus online tuning.}
-Table~\ref{tab:full} shows that the SQF policy
-works quite well on the traces considered here on average, and under this
-resampling scheme. The variability however, is high, and this policy is not the
-best on all traces. It interesting to have an experimental evaluation of
-exactly how much one gains on average by using an online policy versus a
-leave-one-trace-out majority vote policy choice.  However, a good estimation of
-this quantity is tricky. Indeed, we expect the change in the relative
-performance of policies to increase with two factors. First, simply adding
-workload traces to our selection will naturally show more cases where behavior
-is different. Second, factoring in the topology model of the machines should
-increase the heterogeneity between platforms and therefore increase the
-dependence of the simulation performance on the trace/platform. Such a study
-would be much more tedious. Here the need for adaptivity is actually
-understated by the simulation approach.
+Table~\ref{tab:full} shows that the SQF policy works quite well on the traces
+considered here on average, and under this resampling scheme. However, the
+variability is high, and this policy is not the best on all traces. It
+interesting to have an experimental evaluation of exactly how much one gains on
+average by using an online policy versus a leave-one-trace-out majority vote
+policy choice.  However, a good estimation of this quantity is tricky. Indeed,
+we expect the change in the relative performance of policies to increase with
+two factors. First, simply adding workload traces to our selection will
+naturally show more cases with a different behavior. Second, factoring in the
+topology model of the machines should increase the heterogeneity between
+platforms and therefore increase the dependence of the simulation performance
+on the trace/platform. Such a study would be much more ambitious and is beyond
+the scope of this study. Here the need for adaptivity is actually understated
+by the simulation approach.
 
@@
 we can comment on the added complexity and computing overhead of using the
-simulation-based and bandit online strategies.  In both these cases, the
-scheduling decisions do not have an overhead, as the policy is fixed during the
-period (one day, or week). The bandit strategy has a negligible computational
-overhead at period end. Its two computational operations are maintaining a
+simulation-based and bandit online strategies. In both cases, there is no
+overhead in scheduling decisions as the policy is fixed during the period (one
+day, or week). The bandit strategy has a negligible computational overhead at
+the end of the period. Its two computational operations are maintaining a
 sequential average and generating a pseudo-random number.  The simulation-based
@@
 the number of alternative policies considered (here, K=12). Using our
-simulator, This takes a few seconds. Using a more precise simulator with
+simulator, this takes a few seconds. Using a more precise simulator with
 topological modeling, one can argue that the overhead is still manageable since
@@
 is that these simulations should not take too much time compared to the period
-size, as to not delay the policy switch decision too much.
+size, in order not to delay too much the policy switch decision.
 

### 2018-03-09 paper/revision2.tex
fix

@@
 Let us first sincerely thank the reviewers for their relevant and in-depth
-remarks once again. Their comments helped us further the discussion of the
-results, and find a typographical error in the presentation of the results that
-passed our scrutiny.
+remarks once again. Their comments helped to further discuss the results, and
+to find a typographical error in the presentation of the results that passed
+our scrutiny.
 
@@
 
-Regarding your first point, good remark, we based this experimental design
-around previous results from "Tuning Backfilling Queues"(JSSPP17) that show
-some degree of stationarity using a train/test procedure. We made this clearer
-in the text.
-We indeed expect that non-stationary policy adaptation may allow more
-improvements (while simultaneously being harder to achieve).
+Regarding your first point, this is a good remark, we based this experimental
+design around previous results from "Tuning Backfilling Queues"(JSSPP17) that
+show some degree of stationarity using a train/test procedure. We clarified
+this in the text.
+We indeed expect that non-stationary policy adaptation may
+allow more improvements (while simultaneously being much harder to achieve).
 
-Your second point is also relevant, yes that should improve the results.
+Your second point is also relevant: starting the simulation at the end of the previous
+period should improve the results.
 But we can see that the 'full feedback' simulated policy here is almost as good as the best
 policy in the set, so we did not worry about this. In a non-stationary case, it may indeed
-be necessary to properly take care of this.
+be necessary to properly take care of this. We added a remark in section 5.1:
+
+\begin{displayquote}
+The above cost directly corresponds to the cumulative waiting time of the
+policy over the preceding periods (more precisely to the cumulative simulated
+estimate of the waiting time of the policy over the preceding periods) so that
+the length of the period has little impact here. The only bias comes from the
+boundary states of the simulation. Indeed, we run simulations from an empty
+system and wait for the system to be empty when job submissions cease at the
+end of the period (see Figure~\ref{fig:resimu}). Note that starting the simulation
+using the system state at the end of the previous period should improve the results.
+Here it will not be necessary hence we keep the simple version with an empty system.
+However, it may prove useful in the non-stationary case mentionned above.
+\end{displayquote}
 
@@
 results in significant improvements over existing methods.
-We thank you for your in-depth remarks once again. We integrated these comments into the minor revision.
+
+We thank you once again for your in-depth comments which we integrated into the current revision.
 
@@
 
-\textbf{The authors mention "reducing uncertainty" in Section 1.1. It is not clear why this reduction is relevant within the
-manuscript (except as a justification for adding noise?)
-The first paragraph of Section 1.1 is relevant to the content of the manuscript. The study of the manuscript uses old
-traces that are far from extreme scale parallel computing. The relevance to distributed computing is not clear since
-only traces from parallel computers are used.
-}
+%\textbf{The authors mention "reducing uncertainty" in Section 1.1. It is not clear why this reduction is relevant within the
+%manuscript (except as a justification for adding noise?)
+%The first paragraph of Section 1.1 is relevant to the content of the manuscript. The study of the manuscript uses old
+%traces that are far from extreme scale parallel computing. The relevance to distributed computing is not clear since
+%only traces from parallel computers are used.
+%}
 
-\textbf{Why do the authors discuss conservative backfilling in Section 2 although it does not matter in the rest of the
-manuscript?
-What is the meaning of "dependency of scheduling metrics on the workload"? The authors only apply a single scheduling
-metric (average waiting time) and do not discuss such dependency. Why is it important that previous work does not
-address this dependency in the context of this manuscript?
-}
+%\textbf{Why do the authors discuss conservative backfilling in Section 2 although it does not matter in the rest of the
+%manuscript?
+%What is the meaning of "dependency of scheduling metrics on the workload"? The authors only apply a single scheduling
+%metric (average waiting time) and do not discuss such dependency. Why is it important that previous work does not
+%address this dependency in the context of this manuscript?
+%}
 
-\textbf{The authors mention duality between cumulative and maximal scheduling costs. Why is this duality relevant in the
-context of this manuscript? In Section 3.3 the authors claim that "Section 4.2 will outline our approach to address the
-biobjective aspect of this problem". However, Section 4.2. is very brief and does not discuss such aspect. The reviewer
-did not find any discussion in the rest of the manuscript. Altogether not all references cited in the manuscript are
-relevant for the content of the manuscript.
-}
+%\textbf{The authors mention duality between cumulative and maximal scheduling costs. Why is this duality relevant in the
+%context of this manuscript? In Section 3.3 the authors claim that "Section 4.2 will outline our approach to address the
+%biobjective aspect of this problem". However, Section 4.2. is very brief and does not discuss such aspect. The reviewer
+%did not find any discussion in the rest of the manuscript. Altogether not all references cited in the manuscript are
+%relevant for the content of the manuscript.
+%}
 
-\textbf{The design of the study is contradictory: on the one hand, the authors only select a single metric since this metric
-"is one of the more commonly used objectives". On the other hand, they select 12 reordering strategies without giving
-any justification based on the strategy. They simply state that they want the search space "being as semantically
-diverse as possible" without stating that they achieved this goal with the selection of these strategies. Why do they
-need all the 12 strategies and why not using more strategies? Again they only remark that the strategies "include most
-policies from related works that we are aware of." Such justification is hardly sufficient for a scientific study.
-}
+%\textbf{The design of the study is contradictory: on the one hand, the authors only select a single metric since this metric
+%"is one of the more commonly used objectives". On the other hand, they select 12 reordering strategies without giving
+%any justification based on the strategy. They simply state that they want the search space "being as semantically
+%diverse as possible" without stating that they achieved this goal with the selection of these strategies. Why do they
+%need all the 12 strategies and why not using more strategies? Again they only remark that the strategies "include most
+%policies from related works that we are aware of." Such justification is hardly sufficient for a scientific study.
+%}
 
-\textbf{In Section 5, the authors claim that "the distribution of the job submission process does not radically differ between
-consecutive periods." This claim is not true for every period length. For instance, when using 12 hours as period
-length, two consecutive periods may cover day and night with different submission processes.
-}
+%\textbf{In Section 5, the authors claim that "the distribution of the job submission process does not radically differ between
+%consecutive periods." This claim is not true for every period length. For instance, when using 12 hours as period
+%length, two consecutive periods may cover day and night with different submission processes.
+%}
 
-\textbf{The allocation of jobs to periods based on the submission time may be problematic. For instance a long running job may
-be submitted in one period, waiting for execution in the next period and then executing in the next several periods.
-Therefore, it influences several periods. The authors briefly mention such dependencies (Section 5.3) and state that
-there may be a bias without discussing this issue.
-}
+%\textbf{The allocation of jobs to periods based on the submission time may be problematic. For instance a long running job may
+%be submitted in one period, waiting for execution in the next period and then executing in the next several periods.
+%Therefore, it influences several periods. The authors briefly mention such dependencies (Section 5.3) and state that
+%there may be a bias without discussing this issue.
+%}
 
-\textbf{The authors still keep the decaying discount factor $\lambda$ without using it in the study. In an exploratory study, the
-authors can either ignore such factor because it is not relevant (and saying why this claim holds) or they must include
-it into the study.
-}
+%\textbf{The authors still keep the decaying discount factor $\lambda$ without using it in the study. In an exploratory study, the
+%authors can either ignore such factor because it is not relevant (and saying why this claim holds) or they must include
+%it into the study.
+%}
 
-\textbf{The authors introduce a noise factor to deal with imprecision. This way they introduce an approach to model
-imprecision. But they do not show that their model is suitable in reality.
-}
+%\textbf{The authors introduce a noise factor to deal with imprecision. This way they introduce an approach to model
+%imprecision. But they do not show that their model is suitable in reality.
+%}
 
-\textbf{In Section 5.2, the authors claim that the length of the period has little impact here as the cost metric corresponds
-to the cumulative simulated waiting time for each policy. As already mentioned, this may not be true for periods that
-are small compared to the waiting time of jobs.
-}
+%\textbf{In Section 5.2, the authors claim that the length of the period has little impact here as the cost metric corresponds
+%to the cumulative simulated waiting time for each policy. As already mentioned, this may not be true for periods that
+%are small compared to the waiting time of jobs.
+%}
 
-\textbf{Figure 3 does not illustrate the process of the $\epsilon$-greedy strategy since there is no $\epsilon$ in Figure 3.
-}
+%\textbf{Figure 3 does not illustrate the process of the $\epsilon$-greedy strategy since there is no $\epsilon$ in Figure 3.
+%}
 
-\textbf{Section 6.1.1 mentions cores in traces although several of the old traces use machines that do not have multicore
-processors.
-}
+%\textbf{Section 6.1.1 mentions cores in traces although several of the old traces use machines that do not have multicore
+%processors.
+%}
 
-\textbf{The authors use 40 hours as threshold and justify this selection by referring to another publication. The reader does
-not know about the arguments without reading this other publication. Since the design of the study is at the core of
-this manuscript, such justification if not acceptable.
-}
+%\textbf{The authors use 40 hours as threshold and justify this selection by referring to another publication. The reader does
+%not know about the arguments without reading this other publication. Since the design of the study is at the core of
+%this manuscript, such justification if not acceptable.
+%}
 
-\textbf{The authors state that they have simulated a total of 92400 years. However, this absolute figure is not relevant
-without showing the coverage of the problem space.
-}
+%\textbf{The authors state that they have simulated a total of 92400 years. However, this absolute figure is not relevant
+%without showing the coverage of the problem space.
+%}
 
-\textbf{There are only few language problems in the manuscript:
-Software and work are used in the singular form. The authors must adjust the verb correspondingly (Section 1.1)
-wokload in Fig. 2 and 3
-}
+%\textbf{There are only few language problems in the manuscript:
+%Software and work are used in the singular form. The authors must adjust the verb correspondingly (Section 1.1)
+%wokload in Fig. 2 and 3
+%}
 
@@
 \begin{displayquote}
-Table~3 shows that the SQF policy
-works quite well on the traces considered here on average, and under this
-resampling scheme. The variability however, is high, and this policy is not the
-best on all traces. It interesting to have an experimental evaluation of
-exactly how much one gains on average by using an online policy versus a
-leave-one-trace-out majority vote policy choice.  However, a good estimation of
-this quantity is tricky. Indeed, we expect the change in the relative
-performance of policies to increase with two factors. First, simply adding
-workload traces to our selection will naturally show more cases where behavior
-is different. Second, factoring in the topology model of the machines should
-increase the heterogeneity between platforms and therefore increase the
-dependence of the simulation performance on the trace/platform. Such a study
-would be much more tedious. Here the need for adaptivity is actually
-understated by the simulation approach.
+
+Table~3 shows that the SQF policy works quite well on the traces considered
+  here on average, and under this resampling scheme. However, the variability
+  is high, and this policy is not the best on all traces. It interesting to
+  have an experimental evaluation of exactly how much one gains on average by
+  using an online policy versus a leave-one-trace-out majority vote policy
+  choice.  However, a good estimation of this quantity is tricky. Indeed, we
+  expect the change in the relative performance of policies to increase with
+  two factors. First, simply adding workload traces to our selection will
+  naturally show more cases with a different behavior. Second, factoring in the
+  topology model of the machines should increase the heterogeneity between
+  platforms and therefore increase the dependence of the simulation performance
+  on the trace/platform. Such a study would be more ambitious and is beyond the
+  scope of this study. Here the need for adaptivity is actually understated by
+  the simulation approach.
+
 \end{displayquote}
@@
 
-Right, we need to make this clearer. We added this discussion to paragraph
+Right, this point needs a clarification. We added the following discussion to paragraph
 \textbf{SQF versus online tuning} in Section 6.3.2:
@@
 we can comment on the added complexity and computing overhead of using the
-simulation-based and bandit online strategies. In both these cases, the
-scheduling decisions do not have an overhead, as the policy is fixed during the
-period (one day, or week). The bandit strategy has a negligible computational
-overhead at period end. Its two computational operations are maintaining a
+simulation-based and bandit online strategies. In both cases, there is no
+overhead in scheduling decisions as the policy is fixed during the period (one
+day, or week). The bandit strategy has a negligible computational overhead at
+the end of the period. Its two computational operations are maintaining a
 sequential average and generating a pseudo-random number.  The simulation-based
@@
 is that these simulations should not take too much time compared to the period
-size, as to not delay the policy switch decision too much.
+size, in order not to delay too much the policy switch decision.
 \end{displayquote}

### 2018-03-09 paper/paper.tex
added Index term
paper modif

@@
 \begin{document}
+
+\IEEEtitleabstractindextext{D.4.7.f Parallel Systems, D.4.1.e Scheduling, 1.2.6 Learning}
 
@@
 
+\begin{IEEEbiographynophoto}{Eric Gaussier}
+Biography text here.
+\end{IEEEbiographynophoto}
+
+\begin{IEEEbiographynophoto}{J\'er\^ome Lelong}
+Biography text here.
+\end{IEEEbiographynophoto}
+
+\begin{IEEEbiographynophoto}{Valentin Reis}
+Valentin is a PhD candidate at Inria/University Grenoble Alpes, France.
+He is broadly interested in machine learning and functionnal programming.
+\end{IEEEbiographynophoto}
+
+\begin{IEEEbiographynophoto}{Denis Trystram}
+Biography text here.
+\end{IEEEbiographynophoto}
+
 \end{document}

### 2018-03-13 paper/paper.tex
last version.

--- 2018-03-09 paper/paper.tex
@@
 jerome.lelong@imag.fr}  \\
-\IEEEauthorblockN{Valentin Reis, Denis Trystram}\\
+\IEEEauthorblockN{Valentin Reis}\\
+\IEEEauthorblockA{Univ. Grenoble Alpes, Inria, CNRS, Grenoble INP, LIG\\
+valentin.reis@imag.fr}\\
+\IEEEauthorblockN{Denis Trystram,~\IEEEmembership{Senior Member,~IEEE,}}\\
 \IEEEauthorblockA{Univ. Grenoble Alpes, Inria, CNRS, Grenoble INP, LIG\\
@@
 \begin{document}
+
+\IEEEtitleabstractindextext{D.4.7.f Parallel Systems, D.4.1.e Scheduling, I.2.6 Learning}
 
@@
 
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/eric}}]{Eric Gaussier}
+Prof. Eric Gaussier is known for his work on the intersection of Artificial
+Intelligence (AI) and Data Science (DS), in particular for his contributions on
+models and alogrithms to extract information, insights and knowledge from
+texts. He has worked on three main subfields of AI and DS: machine learning,
+information retrieval and computational linguistics. He is also interested in
+modeling how textual information is shared in social content networks, and how
+such networks evolve over time. More recently, he has also been working on
+improving job scheduling techniques through machine learning, and in learning
+representations for different types of sequences, as texts and time series. He
+currently is Director of the Laboratory of Informatics in Grenoble (LIG).
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/jerome}}]{J\'er\^ome Lelong}
+Jérôme Lelong is an associate professor at Grenoble Institute of Technology and
+member of the LJK laboratory. Since 2011, he has been the head of the master
+program on financial engineering. He is a specialist of stochastic
+optimization and HPC algorithms for learning. He has studied many application
+going from quantitative finance to rare event simulation for electronic
+circuit design.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/valentin}}]{Valentin Reis}
+Valentin is a PhD candidate at Inria/University Grenoble Alpes, France.  He is
+currently working on using learning to improve scheduling processes in
+high performance computing. He is broadly interested in machine learning and
+functionnal programming.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/denis}}]{Denis Trystram}
+Denis Trystram is distinguished professor in the institute of technology of
+Univ. Grenoble-Alpes. He is an honorary member of the prestigious Institut
+Universitaire de France. He was vice-director of the LIG in Grenoble for 2
+years (2014-2015) where he is leading a research group on resource management
+for parallel and distributed systems in a joint team with Inria. Since 2010, he
+is director of the international Master program in Computer Science at Univ.
+Grenoble-Alpes. He is a well-known specialist of scheduling in the broad
+variety of parallel and distributed platforms (including HPC clusters, clouds,
+Internet of Things) and multi-objective optimization (in particular with a
+focus on energy). Currently, he is interested in learning and its link with
+resource management.
+\end{IEEEbiography}
+
 \end{document}
